### 2019-06-25 paper draft.Rmd
Initial commit.

---
title: "FATE summary results"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rcpp)
library(ewidata)
library(knitr)
library(reshape2)
library(rstan)
library(tibble)
library(dplyr)
library(gtools)
library(bayesdfa)
library(assertthat)
library(rstanarm)

```

## Results

This is the initial write-up on results for the FATE paper.

### Climate

The best model for GOA climate was the two-trend, independent error model. Trend one positively loaded SST and the Papa advection index, and negatively loaded GAK1 20m salinity. Trend two negatively loaded the upwelling indices and positively loaded the SLP gradient (Fig. 1A).

#### Fig. 1A
GOA climate trend loadings.

```{r, echo=F}

GOA.clim <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_clim_5.15.18.rds") # read in GOA climate model

# recover the years and names for plots!
ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
keep <- grep("AKCLIM_GOA", ewidata$code)

ewidata <- ewidata[keep,]
sub_data = ewidata

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(GOA.clim$best_model)

# set new names
new.names <- c("East.spring.SST", "East.winter.SST", "GAK1.salinity", "SLP.gradient", "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149", "Papa.advection", "West.spr.SST", "West.winter.SST")

pdf("goa.climate.loadings.pdf", 6,6)
print(plot_loadings(rotated, names = new.names, threshold=0.9))
dev.off()
```

### Fig. 1B
GOA climate trends

```{r, echo=F}
# trends
pdf("goa.climate.trends.pdf", 6,6)
print(plot_trends(rotated, years = as.numeric(colnames(Y))))
dev.off()
```

Find swans.

```{r, echo=F}
xx <- find_swans(rotated, threshold = 0.01, plot=F)
xx$year <- 1950:2016
xx <- filter(xx, below_threshold == T)
xx
```

Test for regime probability. Use n=3 maximum states.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(99)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

# repeat with 2nd and third trends, e.g.
set.seed(99)
f2 = find_regimes(y = y[2,], sds = sd_y[2,], max_regimes = 3)

# # repeat with 2nd and third trends, e.g.
# set.seed(99)
# f3 = find_regimes(y = y[3,], sds = sd_y[3,], max_regimes = 3)
```

Regime results for Trend 1. The 2-regime model is best-supported, no evidence at this point for a switch to a novel, warmer state.

Results for the other two trends not shown - no evidence of a change in state for those.

```{r, echo=F}
print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)
```

## GOA biology results.

One trend model was the best.

```{r, echo=F}
GOA.biol <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_biol_5.15.18.rds") 
GOA.biol$summary
```

### Fig. 2A

```{r, echo=F}

# reload the version of data used
# load updated ewidata!
ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))

# limit to GOA
GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]

# drop CI pinks as they aren't present in odd and even years
GOA <- filter(GOA, code != "CI.PI")

# drop climate
drop <- grep("AKCLIM", GOA$code)
GOA <- GOA[-drop,]

# drop dogfish (too many NAs)
drop <- grep("DOGF", GOA$code)
GOA <- GOA[-drop,]

# drop birds and icy strait and chatnam and seward line and CPR
drop <- c(grep("PROD.", GOA$code), grep("PHEN.", GOA$code), grep("ICY", GOA$code), grep("CHAT", GOA$code), grep("SEWARD", GOA$code), 
          grep("mesozoo", GOA$code), grep("diatoms", GOA$code), grep("copepod", GOA$code))
GOA <- GOA[-drop,]

# and start in 1972
GOA <- GOA[GOA$year>=1972,]

# herring time series are repeats, remove
drop <- grep("SE.HER.ROE", GOA$code)
GOA <- GOA[-drop,]

# and drop the ichthyo data
drop <- grep("ICH", GOA$code)
GOA <- GOA[-drop,]

# recover the years and names for plots!
sub_data = GOA

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(GOA.biol$best_model)

new.names <- c("Chignik.coho", "Chignik.pink", "Chiniak.eulachon", "Chiniak.jellyfish", "Chiniak.shrimp", "Cook.Inlet.coho", "Kodiak.coho", "Kodiak.pink", "Pavlof.capelin", "Pavlof.eulachon", "Pavlof.jellyfish", "Pavlof.shrimp", "Prince.William.coho", "Prince.William.pink", "Southeast.coho", "Southeast.herring", "Southeast.pink", "South.Peninsula.coho", "South.Peninsula.pink")

# loadings
pdf("goa.biol.loadings.pdf", 6, 6)
print(plot_loadings(rotated, names = new.names, threshold=0.9))
dev.off()
```

```{r, echo=F}
# trends

pdf("goa.biol.trend.pdf", 6,6)
print(plot_trends(rotated, years = as.numeric(colnames(Y))))
dev.off()
```

Test for regime probability. Use n=3 maximum states.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(99)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)

```

Next step is to examine relationships between SST and the GOA biology trend.

We're interested in testing the hypothesis that the slope on sst was greater during 1972:1988 (high Aleutian Low variance) than during 1989:2013 (low Aleutian Low variance). Also interested in asking which of these periods the 2014-106 heat wave years are most similar to.

```{r, include=F}
# first, define the function
trend_lm <- function(rotated_modelfit,
  all_years, selected_years,
  climate_dat = NULL,
  trend_number = 1, samples = 50,
  ITER = 1000, CHAINS = 1, n_eff = 50, rhat = 1.1, show_plot = TRUE) {

  total_mcmc <- dim(rotated_modelfit$trends)[1]
  draws <- sample(seq_len(total_mcmc), samples)

  yrs_i <- which(all_years %in% selected_years)
  climate_dat <- dplyr::filter(climate_dat, Year %in% selected_years)

  predictor_name = names(climate_dat)[which(names(climate_dat) != "Year")]
  formula_str = paste("dfa_trend_draw ~", predictor_name[1])
  if(length(predictor_name) > 1) {
    for(i in 2:length(predictor_name)) {formula_str = paste(formula_str, predictor_name[i], sep=" + ")}
  }
  formula <- as.formula(formula_str)

  post <- list()
  for (i in seq_along(draws)) {
    dd = climate_dat
    dd$dfa_trend_draw = rotated_modelfit$trends[draws[i], trend_number, yrs_i]
    m_temp <- stan_glm(formula, data = dd,
      iter = ITER, chains = CHAINS)
    assert_that(all(m_temp$stan_summary[,"Rhat"] < rhat))
    assert_that(all(m_temp$stan_summary[,"n_eff"] > n_eff))
    post[[i]] <- as.data.frame(as.matrix(m_temp))
  }
  post_all <- dplyr::bind_rows(post)

  if (show_plot) {
    #par(mfrow = c(1, 2))
    #plot(post_all[,predictor], type = "l", col = "#00000080", lwd = 0.5)
    #abline(v = seq(0, ITER * length(draws), ITER), col = "#FF000050")
    #plot(density(post_all[,predictor]))
    #abline(v = 0, col = "#00000050")
  }
  invisible(post_all)
}

#  load sst as the explanatory variable.

ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
keep <- grep("AKCLIM_GOA", ewidata$code)
clim.var <- ewidata[keep,]  
keep <- grep("win.sst", clim.var$code)
clim.var <- clim.var[keep,]

win.sst <- tapply(clim.var$value, clim.var$year, mean)

years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2016

# set up climate data
cli_dat <- data.frame(Year=as.numeric(names(win.sst)),SST=win.sst)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + SST:slope, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod = trend_lm(rotated_modelfit = r, all_years = 1972:2016,
  selected_years = 1972:2016, climate_dat=cli_dat)

names(mod)[2:4] <- c("1972-1988", "1989-2013", "2014-2016")
summary(mod)

# get 95% CIs

for(i in 2:4){
x <- sort(mod[,i])
print(names(mod)[i])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))
}

melt.mod <- melt(mod, measure.vars=c("1972-1988", "1989-2013", "2014-2016"))
names(melt.mod)[3] <- "era"

pdf("Fig. 3A goa.biol.on.sst.pdf", 6, 4)
ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="A) SST")
dev.off()

```

Now the PDO.

```{r, echo=F}

pdo <- read.csv("pdo.csv")

win.pdo <- pdo$NDJFM

years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2016

# set up climate data
cli_dat <- data.frame(Year=pdo$YEAR,PDO=win.pdo)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(PDO = arm::rescale(PDO))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + PDO:slope, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod2 = trend_lm(rotated_modelfit = r, all_years = 1972:2016,
  selected_years = 1972:2016, climate_dat=cli_dat)

names(mod2)[2:4] <- c("1972-1988", "1989-2013", "2014-2016")

summary(mod2)

# get 95% CIs

for(i in 2:4){
x <- sort(mod2[,i])
print(names(mod2)[i])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))
}

melt.mod <- melt(mod2, measure.vars=c("1972-1988", "1989-2013", "2014-2016"))
names(melt.mod)[3] <- "era"

pdf("Fig. 3B goa.biol.on.pdo.pdf", 6, 4)
ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) PDO")
dev.off()

```

And the NPGO.

```{r, echo=F}

# 
# npgo <- read.csv("winter.npgo.csv")
# 
# years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2016
# 
# # set up climate data
# cli_dat <- data.frame(Year=npgo$YEAR,NPGO=npgo$NDJFM)
# 
# cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(NPGO = arm::rescale(NPGO))
# 
# # cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
# 
# cli_dat$slope <- 1
# cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
# cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
# cli_dat$slope <- as.factor(cli_dat$slope)
# 
# # rename rotated
# r <- rotated
# 
# Y = rnorm(length(cli_dat$slope))
# cli_dat = as.data.frame(model.matrix(Y ~ Year + NPGO:slope, data=cli_dat))
# # replace ':' in name for formula
# names(cli_dat) = gsub(":","",names(cli_dat))
# # remove named intercept
# cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
# 
# mod2 = trend_lm(rotated_modelfit = r, all_years = 1972:2016,
#   selected_years = 1972:2016, climate_dat=cli_dat)
# 
# 
# summary(mod2)
# 
# names(mod2)[2:4] <- c("1972-1988", "1989-2013", "2014-2016")
# 
# melt.mod <- melt(mod2, measure.vars=c("1972-1988", "1989-2013", "2014-2016"))
# names(melt.mod)[3] <- "era"
# 
# pdf("Fig. 3C goa.biol.on.npgo.pdf", 6, 4)
# ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) NPGO")
# dev.off()

```

## GOA ichthyo results.

One trend model was again the best.

```{r, echo=F}
GOA.ich <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_ich_7.19.18.one.trend.rds") 
GOA.ich$summary
```

### Fig. _A

```{r, echo=F}

# reload the version of data used
# load updated ewidata!
# load updated ewidata!
ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))

# limit to GOA
GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
unique(GOA$code)

# limit to ichthy

keep <- grep("ICH", GOA$code)

GOA <- GOA[keep,]

# recover the years and names for plots!
sub_data = GOA

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(GOA.ich$best_model)

new.names <- c("sandlance", "ronquils", "walleye.pollock", "Pacific.cod", "flathead.sole", "n.rock.sole", "rockfish")

# loadings
pdf("goa.ich.loadings.pdf", 6, 6)
print(plot_loadings(rotated, names = new.names, threshold=0.9))
dev.off()
```

```{r, echo=F}
# trends

pdf("goa.ich.trend.pdf", 6,6)
print(plot_trends(rotated, years = as.numeric(colnames(Y))))
dev.off()
```

Test for regime probability. Use n=3 maximum states.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(11)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)

```

Next step is to examine relationships between SST and the ichthyo trend.

We're interested in testing the hypothesis that the slope on sst was greater during 1972:1988 (high Aleutian Low variance) than during 1989:2013 (low Aleutian Low variance). Also interested in asking which of these periods the 2014-106 heat wave years are most similar to.

```{r, include=F}

#  load sst as the explanatory variable.

ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
keep <- grep("AKCLIM_GOA", ewidata$code)
clim.var <- ewidata[keep,]  
keep <- grep("win.sst", clim.var$code)
clim.var <- clim.var[keep,]

win.sst <- tapply(clim.var$value, clim.var$year, mean)

years <- as.numeric(colnames(GOA.ich$best_model$data))

# set up climate data
cli_dat <- data.frame(Year=as.numeric(names(win.sst)),SST=win.sst)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

# this time series is too short to divide up into eras!
cli_dat$slope <- 1
#cli_dat$slope[cli_dat$Year %in% 1997:2013] <- 2
#cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + SST, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod = trend_lm(rotated_modelfit = r, all_years = years,
  selected_years = years, climate_dat=cli_dat)

summary(mod)

# get 95% CIs
x <- sort(mod[,2])
print(names(mod)[2])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))

# dropping this b/c we have no eras for this short TS!
# melt.mod <- melt(mod, measure.vars=c("1972-1988", "1989-2013", "2014-2016"))
# names(melt.mod)[3] <- "era"

pdf("Fig. x goa.ich.on.sst.pdf", 4, 5)
ggplot(mod, aes(x=SST)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + xlab("Slope")
dev.off()

```

Now the zoops model.

```{r, echo=F}
GOA.zoop <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_zoop_7.19.18.one.trend.rds") 
GOA.zoop$summary
```

```{r, echo=F}

# load updated ewidata!
ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))

# limit to GOA
GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
unique(GOA$code)

# drop salmon
drop <- grep("PI", GOA$code)
GOA <- GOA[-drop,]

drop <- grep("CO", GOA$code)
GOA <- GOA[-drop,]

# drop climate
drop <- grep("AKCLIM", GOA$code)
GOA <- GOA[-drop,]

unique(GOA$code)

# drop Pavlof and Chiniak and herring and kittiwakes
drop <- grep("PAV", GOA$code)
GOA <- GOA[-drop,]

drop <- grep("CHI", GOA$code)
GOA <- GOA[-drop,]

drop <- grep("HER", GOA$code)
GOA <- GOA[-drop,]

drop <- grep("BLKI", GOA$code)
GOA <- GOA[-drop,]

# drop total icy zoops!
drop <- grep("ICY.ZOOP.TOTDEN", GOA$code)
GOA <- GOA[-drop,]

# and drop ichthyoplankton
drop <- grep("ICH", GOA$code)
GOA <- GOA[-drop,]

unique(GOA$code)
# recover the years and names for plots!
sub_data = GOA

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(GOA.zoop$best_model)

new.names <- c("CPR.copepods", "CPR.diatoms", "Icy.Str.amphipods", "Icy.Str.euphausids", "Icy.Str.gastropods", "CPR.mesozooplankton", "Seward.A.calanoid", "Seward.A.euphausid", "Seward.B.calanoid", "Seward.B.euphausid", "Seward.LA.euphausid", "Seward.LB.calanoid", "Seward.LB.euphausid", "Seward.calanoid.size", "Seward.euphausid.size", "Seward.LA.calanoid")

# loadings
pdf("goa.zoop.loadings.pdf", 6, 6)
print(plot_loadings(rotated, names = new.names, threshold=0.9))
dev.off()
```

```{r, echo=F}
# trends

pdf("goa.zoop.trend.pdf", 6,6)
print(plot_trends(rotated, years = as.numeric(colnames(Y))))
dev.off()
```

Test for regime probability. Use n=3 maximum states.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(11)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)

```

Next step is to examine relationships between SST and the zoop trend.

We're interested in testing the hypothesis that the slope on sst was greater during 1972:1988 (high Aleutian Low variance) than during 1989:2013 (low Aleutian Low variance). Also interested in asking which of these periods the 2014-106 heat wave years are most similar to.

```{r, include=F}

#  load sst as the explanatory variable.

ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
keep <- grep("AKCLIM_GOA", ewidata$code)
clim.var <- ewidata[keep,]  
keep <- grep("win.sst", clim.var$code)
clim.var <- clim.var[keep,]

win.sst <- tapply(clim.var$value, clim.var$year, mean)

years <- as.numeric(colnames(GOA.zoop$best_model$data))

# set up climate data
cli_dat <- data.frame(Year=as.numeric(names(win.sst)),SST=win.sst)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
#cli_dat$slope[cli_dat$Year %in% 1997:2013] <- 2
#cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + SST, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod = trend_lm(rotated_modelfit = r, all_years = years,
  selected_years = years, climate_dat=cli_dat)

summary(mod)

# get 95% CIs
x <- sort(mod[,2])
print(names(mod)[2])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))

pdf("Fig. x goa.zoop.on.sst.pdf", 4, 5)
ggplot(mod, aes(x=SST)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + xlab("Slope")
dev.off()

```

# Now the EBS results.

```{r, echo=F}

EBS.clim <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/EBS_clim_6.19.18.rds") # read in EBS climate model

EBS.clim$summary
```

And the trends and loading plots.

```{r, echo=F}

# recover the years and names for plots!
ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
keep <- grep("AKCLIM_EBS", ewidata$code)

ewidata <- ewidata[keep,]
sub_data = ewidata

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(EBS.clim$best_model)

new.names <- c("Bottom.temp", "Spring.SST", "Winter.SST", "Summer.NW.wind", "Summer.SE.wind", "Winter.NW.wind", "Winter.SE.wind", "Ice.57.75N", "Ice.58.75N", "Ice.64.25N")

# loadings
pdf("ebs.clim.loadings.pdf", 6, 6)
print(plot_loadings(rotated, names = new.names, threshold=0.9))
dev.off()

# trends
pdf("ebs.clim.trend.pdf", 6,6)
print(plot_trends(rotated, years = as.numeric(colnames(Y))))
dev.off()
```

Find swans.

```{r, echo=F}
xx <- find_swans(rotated, threshold = 0.01, plot=F)
xx$year <- 1950:2016
xx <- filter(xx, below_threshold == T)
xx
```

Test for regime probability. Use n=3 maximum states to rule out a switch to a warmer state.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(99)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

```

Regime results for Trend 1. The 2-regime model is best-supported.

```{r, echo=F}
print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)
```

## EBS biology results.

```{r, echo=F}
EBS.biol <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/EBS_biol_6.19.18.rds") 
EBS.biol$summary
```

Now plot trends and loadings.

```{r, echo=F}

# reload the version of data used
# load updated ewidata!
ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))

# first, EBS biology
codes <- unique(ewidata$code)

# limit to EBS
EBS <- ewidata[ewidata$system == "EBS",]

# and start in 1975, when data beyond salmon are first available
EBS <- EBS[EBS$year>=1975,]

unique(EBS$code)

# drop clim
drop <- grep("AKCLIM", EBS$code)
EBS <- EBS[-drop,]

# drop depth
drop <- grep("DEPTH", EBS$code)
EBS <- EBS[-drop,]

# recover the years and names for plots!
sub_data = EBS

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(EBS.biol$best_model)

new.names <- c("Alaska.place.lat", "Bristol.Bay.coho", "Bering.flounder.lat", "St.George.kitt.phen", "St.Paul.kitt.phen", "St.Paul.kitt.product", "com.eider.eggs/nest", "com.eider.hatch.date", "com.eider.success", "St.George.com.murre.phenol", "St.Paul.com.murre.phenol", "St.George.com.murre.product", "St.Paul.com.murre.product", "juv.cod.cpue", "juv.pollock.cpue", "flathead.sole.lat", "gl.gull.eggs/nest", "gl.gull.hatch.date", "gl.gull.success", "Grnlnd.turbot.lat", "Pac.cod.lat", "Pac.halibut.lat", "red.king.crab.lat", "rock.sole.lat", "shortf.eelpout.lat", "snow.crab.lat", "spec.eider.eggs/nest", "spec.eider.hatch.date", "spec.eider.success", "sturgeon.poacher.lat", "Tanner.crab.lat", "St.George.tb.murre.phenol", "St.Paul.tb.murre.phenol", "St.George.tb.murre.product", "St.Paul.tb.murre.product", "walleye.pollock.lat", "wattled.eelpout.lat", "yellowfin.sole.lat")

# loadings
pdf("ebs.biol.loadings.pdf", 6, 6)
print(plot_loadings(rotated, names = new.names, threshold=0.9))
dev.off()

```

```{r, echo=F}
# trends

pdf("ebs.biol.trend.pdf", 6,6)
print(plot_trends(rotated, years = as.numeric(colnames(Y))))
dev.off()
```
Test for regime probability. Use n=3 maximum states to rule out a switch to a warmer state.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(99)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

```

Regime results for Trend 1. The 2-regime model is best-supported.

```{r, echo=F}
print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)
```

Next step is to examine relationships between SST and the EBS biology trend.

```{r, message=FALSE, results='hide', echo=F}

# load sst as the explanatory variable.

ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
keep <- grep("AKCLIM_EBS", ewidata$code)
clim.var <- ewidata[keep,]  
keep <- grep("win.sst", clim.var$code)
clim.var <- clim.var[keep,]

win.sst <- tapply(clim.var$value, clim.var$year, identity)

years <- as.numeric(colnames(EBS.biol$best_model$data)) # these are the years for the biology model - 1975:2016

cli_dat <- data.frame(Year=as.numeric(names(win.sst)), SST=win.sst)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + SST:slope, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(4)
mod = trend_lm(rotated_modelfit = r, all_years = 1975:2016,
  selected_years = 1975:2016, climate_dat=cli_dat)

names(mod)[2:4] <- c("1975-1988", "1989-2013", "2014-2016")
summary(mod)

# get 95% CIs

for(i in 2:4){
x <- sort(mod[,i])
print(names(mod)[i])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))
}

melt.mod <- melt(mod, measure.vars=c("1975-1988", "1989-2013", "2014-2016"))
names(melt.mod)[3] <- "era"

pdf("Fig. x ebs.biol.on.sst.pdf", 6, 4)
ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="A) SST")
dev.off()
```

Now the PDO.

```{r, message=FALSE, results='hide', echo=F}

pdo <- read.csv("pdo.csv")

win.pdo <- pdo$NDJFM

years <- as.numeric(colnames(EBS.biol$best_model$data)) # these are the years for the biology model - 1972:2016

# set up climate data
cli_dat <- data.frame(Year=pdo$YEAR,PDO=win.pdo)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(PDO = arm::rescale(PDO))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + PDO:slope, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod2 = trend_lm(rotated_modelfit = r, all_years = 1975:2016,
  selected_years = 1975:2016, climate_dat=cli_dat)

names(mod2)[2:4] <- c("1975-1988", "1989-2013", "2014-2016")

summary(mod2)

# get 95% CIs

for(i in 2:4){
x <- sort(mod2[,i])
print(names(mod2)[i])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))
}

melt.mod <- melt(mod2, measure.vars=c("1975-1988", "1989-2013", "2014-2016"))
names(melt.mod)[3] <- "era"

pdf("Fig. x ebs.biol.on.pdo.pdf", 6, 4)
ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) PDO")
dev.off()

```

And the winter AO.

```{r, echo=F}

ao <- read.csv("winter.ao.csv")

win.ao <- ao$NDJFM

years <- as.numeric(colnames(EBS.biol$best_model$data)) # these are the years for the biology model - 1975:2016

# set up climate data
cli_dat <- data.frame(Year=ao$YEAR,AO=win.ao)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(AO = arm::rescale(AO))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + AO:slope, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod3 = trend_lm(rotated_modelfit = r, all_years = 1975:2016,
  selected_years = 1975:2016, climate_dat=cli_dat)

names(mod3)[2:4] <- c("1975-1988", "1989-2013", "2014-2016")

summary(mod3)

# get 95% CIs

for(i in 2:4){
x <- sort(mod3[,i])
print(names(mod3)[i])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))
}

# and proportion > 0
for(i in 2:4){
x <- sort(mod3[,i])
print(names(mod3)[i])
print(paste("prop > 0: ", sum(x > 0)/length(x)))
}

melt.mod <- melt(mod3, measure.vars=c("1975-1988", "1989-2013", "2014-2016"))
names(melt.mod)[3] <- "era"

pdf("Fig. x ebs.biol.on.ao.pdf", 6, 4)
ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) AO")
dev.off()
```

```{r}
head(melt.mod)
restricted <- filter(melt.mod, era %in% c("1975-1988", "1989-2013"))
restricted <- melt.mod
png("ebs community response to ao.png", 6,4, units="in", res=300)
ggplot(data=restricted, aes(fill=era, x=era, y=value)) + geom_violin() + coord_flip() + ggtitle("Community response to Arctic Oscillation") +
  geom_hline(yintercept = 0) + xlab("Era") + ylab("Regression coefficient") + theme_gray() + theme(legend.position = c(0.8, 0.5), legend.title = element_blank())
dev.off()
```

### 2019-09-02 ewidata update - GOA.R
Finished goa climate modeling, uploaded updated data.

@@ -26,12 +26,12 @@ ggplot(goa.dat, aes(year, value)) +
 # 20º-70ºN, 120º-250ºE, 1854-present
 
 # identify latest year and month needed
-# year <- 2019
-# month <- "07"
-# 
-# URL <- paste("https://coastwatch.pfeg.noaa.gov/erddap/griddap/nceiErsstv5.nc?sst[(1854-01-01):1:(", year, "-", month, "-01T00:00:00Z)][(0.0):1:(0.0)][(20):1:(70)][(120):1:(250)]", sep="")
-# 
-# download.file(URL, "North.Pacific.ersst")
+year <- 2019
+month <- "07"
+
+URL <- paste("https://coastwatch.pfeg.noaa.gov/erddap/griddap/nceiErsstv5.nc?sst[(1854-01-01):1:(", year, "-", month, "-01T00:00:00Z)][(0.0):1:(0.0)][(20):1:(70)][(120):1:(250)]", sep="")
+
+download.file(URL, "North.Pacific.ersst")
 nc <- nc_open("North.Pacific.ersst")
 
 # extract dates
@@ -311,6 +311,17 @@ temp <- goa.dat %>%
 
 goa.clim$papa.index <- c(temp$value, papa$papa[3:5])  
 
+plot.clim <- goa.clim %>%
+  gather(key="code", value="value", -year)
+
+ggplot(plot.clim, aes(year, value)) +
+  theme_bw() +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~code, scale="free_y")
+
+write.csv(goa.clim, "updated goa climate data.csv", row.names=F)
+
 #######################################
 # now run the dang climate DFA model! #
 #######################################
@@ -328,31 +339,28 @@ library(gtools)
 library(bayesdfa)
 
 mcmc_iter = 4000
-max_trends = 3 # changing to 3!
+max_trends = 3 
 mcmc_chains = 3
 rstan_options(auto_write = TRUE)
 options(mc.cores = parallel::detectCores())
 
-
+# doing this in a clunky way - 
+# find dfa trends with 2 or 3 candidate trends 
+# and original data (as in previous iteration of paper)
+# or expanded data (adding SSH and wind stress)
 sub_data <- goa.clim %>%
   gather(key="code", value="value", -year)
 
 # set the name for the model output!
-name <- "GOA_clim_3_trends"
+name <- "GOA_clim_3_trends_expanded_data"
 
-
-# this is the DFA code chunk!
-
[…]

### 2019-09-02 ewidata update - GOA.R
Finished goa climate modeling, uploaded updated data.

@@ -26,12 +26,12 @@ ggplot(goa.dat, aes(year, value)) +
 # 20º-70ºN, 120º-250ºE, 1854-present
 
 # identify latest year and month needed
-# year <- 2019
-# month <- "07"
-# 
-# URL <- paste("https://coastwatch.pfeg.noaa.gov/erddap/griddap/nceiErsstv5.nc?sst[(1854-01-01):1:(", year, "-", month, "-01T00:00:00Z)][(0.0):1:(0.0)][(20):1:(70)][(120):1:(250)]", sep="")
-# 
-# download.file(URL, "North.Pacific.ersst")
+year <- 2019
+month <- "07"
+
+URL <- paste("https://coastwatch.pfeg.noaa.gov/erddap/griddap/nceiErsstv5.nc?sst[(1854-01-01):1:(", year, "-", month, "-01T00:00:00Z)][(0.0):1:(0.0)][(20):1:(70)][(120):1:(250)]", sep="")
+
+download.file(URL, "North.Pacific.ersst")
 nc <- nc_open("North.Pacific.ersst")
 
 # extract dates
@@ -311,6 +311,17 @@ temp <- goa.dat %>%
 
 goa.clim$papa.index <- c(temp$value, papa$papa[3:5])  
 
+plot.clim <- goa.clim %>%
+  gather(key="code", value="value", -year)
+
+ggplot(plot.clim, aes(year, value)) +
+  theme_bw() +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~code, scale="free_y")
+
+write.csv(goa.clim, "updated goa climate data.csv", row.names=F)
+
 #######################################
 # now run the dang climate DFA model! #
 #######################################
@@ -328,31 +339,28 @@ library(gtools)
 library(bayesdfa)
 
 mcmc_iter = 4000
-max_trends = 3 # changing to 3!
+max_trends = 3 
 mcmc_chains = 3
 rstan_options(auto_write = TRUE)
 options(mc.cores = parallel::detectCores())
 
-
+# doing this in a clunky way - 
+# find dfa trends with 2 or 3 candidate trends 
+# and original data (as in previous iteration of paper)
+# or expanded data (adding SSH and wind stress)
 sub_data <- goa.clim %>%
   gather(key="code", value="value", -year)
 
 # set the name for the model output!
-name <- "GOA_clim_3_trends"
+name <- "GOA_clim_3_trends_expanded_data"
 
-
-# this is the DFA code chunk!
-
[…]

### 2019-09-04 ewidata update - GOA biology.R
Updating and processing GOA biology data.

@@ -1,0 +1,40 @@
+library(tidyverse)
+library(ncdf4)
+library(chron)
+library(fields)
+library(maps)
+library(mapdata)
+library(zoo)
+
+dat <- read.csv("ewidata.csv", row.names = 1)
+
+levels(dat$system)
+
+# restrict to GOA 
+
+goa.dat <- dat %>%
+  filter(system %in% c("EGOA", "WGOA"))
+
+# drop climate data!
+drop <- grep("AKC", goa.dat$code)
+goa.dat <- goa.dat[-drop,]
+
+ggplot(goa.dat, aes(year, value)) +
+  theme_bw() +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~code, scales="free_y") 
+
+# examine salmon timeseries alone
+
+salmon <- c(grep(".PI", goa.dat$code), grep(".CO", goa.dat$code))
+
+salm.dat <- goa.dat[salmon,]
+
+ggplot(salm.dat, aes(year, value)) +
+  theme_bw() +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~code, scales="free_y") 
+
+# update

### 2019-09-04 paper draft.Rmd
Further to climate plots and rolling correlations.

@@
 
-GOA.clim <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_clim_5.15.18.rds") # read in GOA climate model
+GOA.clim <- readRDS("GOA_clim_2_trends_original_data.rds") # read in GOA climate model
 
 # recover the years and names for plots!
-ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
-keep <- grep("AKCLIM_GOA", ewidata$code)
-
-ewidata <- ewidata[keep,]
-sub_data = ewidata
+goa.clim <- read.csv("updated goa climate data.csv")
+
+sub_data <- goa.clim %>%
+  select(-ssh, -wind.stress) %>%
+  gather(key="code", value="value", -year)
 
@@
 # set new names
-new.names <- c("East.spring.SST", "East.winter.SST", "GAK1.salinity", "SLP.gradient", "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149", "Papa.advection", "West.spr.SST", "West.winter.SST")
+new.names <- c("East.spring.SST", "East.winter.SST", "SLP.gradient", "Papa.advection", "GAK1.salinity", "West.spring.SST", "West.winter.SST", "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149")
+
+# bespoke plotting code
+loadings.1 <- as.data.frame(rotated$Z_rot[,,1])
+loadings.2 <- as.data.frame(rotated$Z_rot[,,2])
+names(loadings.1) <- names(loadings.2) <- names
+
+# drop columns that don't meet threshold
+thresh <- 0.9
+
+f <- function(x) max(sum(x<0)/length(x), sum(x>0)/length(x))
+
+prop1 <- apply(loadings.1, 2, f)
+prop2 <- apply(loadings.2, 2, f)
+
+# so! only Trend 1 includes loadings > 0.9! 
+# this justifies fitting/retaining only a 1-trend model!
+
+# load the 1-trend object
+GOA.clim <- readRDS("GOA_clim_1_trend_original_data.rds") # read in GOA climate model
+
+# recover the years and names for plots!
+goa.clim <- read.csv("updated goa climate data.csv")
+
+sub_data <- goa.clim %>%
+  select(-ssh, -wind.stress) %>%
+  gather(key="code", value="value", -year)
+
+melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
+Y <- dcast(melted, code ~ year)
+names = Y$code
+Y = as.matrix(Y[,-which(names(Y) == "code")])
+
+rotated = rotate_trends(GOA.clim$best_model)
+
+# set new names
+new.names <- c("East.spring.SST", "East.winter.SST", "SLP.gradient", "Papa.advection", "GAK1.salinity", "West.spring.SST", "West.winter.SST", "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149")
 
 pdf("goa.climate.loadings.pdf", 6,6)
-print(plot_loadings(rotated, names = new.names, threshold=0.9))
-dev.off()
+print(plot_loadings(rotated, names = new.names, threshold=0.8))
+dev.off()
+
+plot.trend <- data.frame(year=1950:2019, trend=as.vector(rotated$trends_mean),
+                         lo=as.vector(rotated$trends_lower),
+                         hi=as.vector(rotated$trends_upper))
+
+plot.b <- ggplot(plot.trend, aes(year, trend)) +
+  theme_bw() +
+  geom_line(color=cb[3]) +
+  geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
+  geom_hline(yintercept = 0) + ylab("Trend value") +
+  theme(axis.title.x = element_blank()) + 
+  ggtitle("b) Trend")
+
+png("combined GOA climate loading and trend plot.png", 6, 3, units = "in", res=300)
+
+ggpubr::ggarrange(plot.a, plot.b, ncol=2)
+
+dev.off()
+
+# back to the bespoke code!
+loadings.1 <- as.data.frame(rotated$Z_rot[,,1])
+names(loadings.1)  <- new.names
+
+# drop columns that don't meet threshold
+# first, figure out the proportion above/below 0
+f <- function(x) max(sum(x<0)/length(x), sum(x>0)/length(x))
+
+prop1 <- apply(loadings.1, 2, f)
+prop1
+keep <- prop1 > 0.95
+
+loadings.1 <- loadings.1[,keep]
+
+# plot
+plot.load <- loadings.1 %>%
+  gather()
+
+rank <- tapply(plot.load$value, plot.load$key, mean)
+plot.load$rank <- rank[match(plot.load$key, names(rank))]
+
+plot.load$key <- reorder(plot.load$key, plot.load$rank)
+
+# set pallette
+cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+plot.a <- ggplot(plot.load, aes(key, value)) +
+  theme_bw() +
+  geom_violin(fill=cb[3]) +
+  geom_hline(yintercept = 0) +
+  theme(axis.title.y = element_blank()) + ylab("Loading") +
+  coord_flip() +
+  ggtitle("a) Loadings")
+
+modelfit <- GOA.clim$best_model
+
+  n_ts <- dim(modelfit$data)[1]
+  n_years <- dim(modelfit$data)[2]
+  pred <- predicted(modelfit)
+   res.df <- data.frame(ID = rep(seq_len(n_ts),
+                                 n_years),
+                        Time = sort(rep(seq_len(n_years), n_ts)), 
+                        mean = c(t(apply(pred, c(3, 4), mean))), 
+                        lo = c(t(apply(pred, c(3, 4), quantile, 0.025))),
+                        hi = c(t(apply(pred, c(3, 4), quantile, 0.975))), 
+                        y = c(modelfit$data))
+
+        res.df$ID <- new.names
+
+  res.df$year <-  rep(1950:2019, each=n_ts)
+
+  res.df$mean.res <- res.df$y - res.df$mean
+  res.df$lo.res <- res.df$y - res.df$lo
+  res.df$hi.res <- res.df$y - res.df$hi
+
+ plot.these <- names(keep)[keep]
+
+ res.df <- res.df %>%
+   filter(ID %in% plot.these)
+
+  ggplot(res.df, aes_string(x = "year", y = "mean.res")) + geom_ribbon(aes_string(ymin = "lo.res", 
+        ymax = "hi.res"), fill=cb[3], alpha = 0.4) + geom_line(color=cb[3]) + facet_wrap("ID", 
+        scales = "free_y") + xlab("Year") + ylab("") + geom_hline(yintercept = 0, col=cb[2]) +
+        geom_vline(xintercept = c(1988.5, 2014), lty=2)
+
+# acf plot
+    index <- unique(res.df$ID)
+
+    par(mfrow=c(2,4))
+    for(i in 1:length(index)){
+     # i <- 1
+      temp <- res.df %>%
+        filter(ID==index[i])
+      temp <- na.omit(temp)
+      acf(temp$mean.res, main=index[i]) 
+    }
+
+# so...no autocorrelation concerns with residuals
+
+# look at rolling correlations between fitted and observed values, as in Ecology paper
+# using 25-yr rolling windows!
+
+    dfa.cor <- data.frame()
+
+    for(i in 1:length(index)){    
+      # i <- 1 
+      temp <- res.df %>%
+        filter(ID==index[i])
+
+       for(ii in 13:58){
+       #  ii <- 13
+          win <- temp[(ii-12):(ii+12),]
+          temp.cor <- data.frame(ID=unique(win$ID),
+                                 year=win$year[13], 
+                                 window=25,
+                                 mean.cor=cor(win$y, win$mean, use="p"),
+                                 n.cor=sum(!is.na(win$y)))
+          dfa.cor <- rbind(dfa.cor, temp.cor)
+       }  }
+
+# now the 11-yr windows
+     for(i in 1:length(index)){    
+      # i <- 1 
+      temp <- res.df %>%
+        filter(ID==index[i])
+             for(ii in 6:65){
+       #  ii <- 13
+          win <- temp[(ii-5):(ii+5),]
+          temp.cor <- data.frame(ID=unique(win$ID),
+                                 year=win$year[6],
+                                 window=11,
+                                 mean.cor=cor(win$y, win$mean, use="p"),
+                                 n.cor=sum(!is.na(win$y)))
+          dfa.cor <- rbind(dfa.cor, temp.cor)
+             }  }
+
+      # and the 15-yr windows
+    for(i in 1:length(index)){    
+      # i <- 1 
+      temp <- res.df %>%
+        filter(ID==index[i])
+
+             for(ii in 8:63){
+       #  ii <- 13
+          win <- temp[(ii-7):(ii+7),]
+          temp.cor <- data.frame(ID=unique(win$ID),
+                                 year=win$year[8],
+                                 window=15,
+                                 mean.cor=cor(win$y, win$mean, use="p"),
+                                 n.cor=sum(!is.na(win$y)))
+          dfa.cor <- rbind(dfa.cor, temp.cor)
+       }  }
+
+# remove salinity correlations with < 10 obs
+dfa.cor <- dfa.cor %>%
+  filter(n.cor>=10)
+ dfa.cor$window <- as.factor(dfa.cor$window)
+
+ggplot(dfa.cor, aes(year, mean.cor, color=window)) + 
+  theme_bw() + theme(axis.title.x = element_blank()) +
+  geom_line() + facet_wrap(~ID, 
+        scales = "free_y") + xlab("Year") + ylab("Correlation")
+
+ggsave("goa climate dfa correlations 11 15 25 year windows.png", height = 4, width=6, units="in")
+
+# and plot 15-yr windows
+ggplot(filter(dfa.cor, window==15), aes(year, mean.cor)) + 
+  theme_bw() + theme(axis.title.x = element_blank()) +
+  geom_line(color=cb[3]) + facet_wrap(~ID, 
+        scales = "free_y") + xlab("Year") + ylab("Correlation")
+
+ggsave("goa climate dfa correlations 15 year windows.png", height = 4, width=6, units="in")
+
+# and plot 25-yr windows
+ggplot(filter(dfa.cor, window==25), aes(year, mean.cor)) + 
+  theme_bw() + theme(axis.title.x = element_blank()) +
+  geom_line(color=cb[3]) + facet_wrap(~ID, 
+        scales = "free_y") + xlab("Year") + ylab("Correlation")
+
+ggsave("goa climate dfa correlations 25 year windows.png", height = 4, width=6, units="in")
+
 ```
@@
 # trends
-pdf("goa.climate.trends.pdf", 6,6)
-print(plot_trends(rotated, years = as.numeric(colnames(Y))))
-dev.off()
+# pdf("goa.climate.trends.pdf", 6,6)
+# print(plot_trends(rotated, years = as.numeric(colnames(Y))))
+# dev.off()
 ```
@@
 xx <- find_swans(rotated, threshold = 0.01, plot=F)
-xx$year <- 1950:2016
+xx$year <- 1950:2019
 xx <- filter(xx, below_threshold == T)
@@
 
-# repeat with 2nd and third trends, e.g.
-set.seed(99)
-f2 = find_regimes(y = y[2,], sds = sd_y[2,], max_regimes = 3)
-
 # # repeat with 2nd and third trends, e.g.
 # set.seed(99)
+# f2 = find_regimes(y = y[2,], sds = sd_y[2,], max_regimes = 3)
+# 
+# # # repeat with 2nd and third trends, e.g.
+# set.seed(99)
 # f3 = find_regimes(y = y[3,], sds = sd_y[3,], max_regimes = 3)
@@
 Regime results for Trend 1. The 2-regime model is best-supported, no evidence at this point for a switch to a novel, warmer state.
-
-Results for the other two trends not shown - no evidence of a change in state for those.
 

### 2019-09-04 ewidata update - GOA.R

@@ -397,9 +397,7 @@ capture.output(summary_table, file = paste0(name, "_summary.txt"))
 
 dev.off() 
 
-
 # now two trends with expanded data
-
 max_trends = 2
 name <- "GOA_clim_2_trends_expanded_data"
 
@@ -452,6 +450,56 @@ sub_data <- goa.clim %>%
   select(-ssh, -wind.stress) %>%
   gather(key="code", value="value", -year)
 
+# ONE! trend model
+# with original data
+max_trends = 1
+name <- "GOA_clim_1_trend_original_data"
+
+# reshape data
+melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
+Y <- dcast(melted, code ~ year)
+names = Y$code
+Y = as.matrix(Y[,-which(names(Y) == "code")])
+
+# do the trend search
+set.seed(99)
+dfa_summary = find_dfa_trends(
+  y = Y,
+  kmax = min(max_trends, nrow(Y)),
+  iter = mcmc_iter,
+  compare_normal = FALSE,
+  variance = c("unequal", "equal"),
+  chains = mcmc_chains
+)
+saveRDS(dfa_summary, file = paste0(name, ".rds"))
+
+# Make default plots (currently work in progress)
+pdf(paste0(name, "_plots.pdf"))
+rotated = rotate_trends(dfa_summary$best_model)
+# trends
+print(plot_trends(rotated, years = as.numeric(colnames(Y))))
+# loadings
+print(plot_loadings(rotated, names = names))
+
+if(ncol(rotated$Z_rot_mean)==2) {
+  plot(rotated$Z_rot_mean[,1], rotated$Z_rot_mean[,2], col="white",
+       xlab="Loading 1", ylab = "Loading 2")
+  text(rotated$Z_rot_mean[,1], rotated$Z_rot_mean[,2], names, cex=0.3)
+  lines(c(-10,10),c(0,0))
+  lines(c(0,0), c(-10,10))
+}
+# predicted values with data
+print(plot_fitted(dfa_summary$best_model,names=names) + 
+        theme(strip.text.x = element_text(size = 6)))
+
+# table of AIC and std errors
+summary_table<-dfa_summary$summary
+capture.output(summary_table, file = paste0(name, "_summary.txt"))
+
+dev.off() 
+
+
+######################
 # two trends with original data
 max_trends = 2
 name <- "GOA_clim_2_trends_original_data"

### 2019-09-04 paper draft.Rmd
Updating GOA biology plots

--- 2019-06-25 paper draft.Rmd
@@
 ```{r, echo=F}
-GOA.biol <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_biol_5.15.18.rds") 
+GOA.biol <- readRDS("GOA_biol_9.4.19.rds") 
 GOA.biol$summary
@@
 # reload the version of data used
-# load updated ewidata!
-ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
-
-# limit to GOA
-GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
-
-# drop CI pinks as they aren't present in odd and even years
-GOA <- filter(GOA, code != "CI.PI")
-
-# drop climate
-drop <- grep("AKCLIM", GOA$code)
-GOA <- GOA[-drop,]
-
-# drop dogfish (too many NAs)
-drop <- grep("DOGF", GOA$code)
-GOA <- GOA[-drop,]
-
-# drop birds and icy strait and chatnam and seward line and CPR
-drop <- c(grep("PROD.", GOA$code), grep("PHEN.", GOA$code), grep("ICY", GOA$code), grep("CHAT", GOA$code), grep("SEWARD", GOA$code), 
-          grep("mesozoo", GOA$code), grep("diatoms", GOA$code), grep("copepod", GOA$code))
-GOA <- GOA[-drop,]
-
-# and start in 1972
-GOA <- GOA[GOA$year>=1972,]
-
-# herring time series are repeats, remove
-drop <- grep("SE.HER.ROE", GOA$code)
-GOA <- GOA[-drop,]
-
-# and drop the ichthyo data
-drop <- grep("ICH", GOA$code)
-GOA <- GOA[-drop,]
+GOA <- read.csv("updated GOA biology data.csv")
 
@@
 
-new.names <- c("Chignik.coho", "Chignik.pink", "Chiniak.eulachon", "Chiniak.jellyfish", "Chiniak.shrimp", "Cook.Inlet.coho", "Kodiak.coho", "Kodiak.pink", "Pavlof.capelin", "Pavlof.eulachon", "Pavlof.jellyfish", "Pavlof.shrimp", "Prince.William.coho", "Prince.William.pink", "Southeast.coho", "Southeast.herring", "Southeast.pink", "South.Peninsula.coho", "South.Peninsula.pink")
+new.names <- c("Chignik.coho", "Chignik.pink", "Chiniak.eulachon", "Chiniak.jellyfish", "Chiniak.shrimp", "Cook.Inlet.coho", "Kodiak.coho", "Kodiak.pink", "Pavlof.capelin", "Pavlof.eulachon", "Pavlof.jellyfish", "Pavlof.shrimp", "Prince.William.coho", "Southeast.coho", "Southeast.herring", "Southeast.pink", "South.Peninsula.coho", "South.Peninsula.pink")
 

### 2019-09-04 ewidata update - GOA biology.R
Updated biology data

@@ -6,35 +6,199 @@ library(maps)
 library(mapdata)
 library(zoo)
 
-dat <- read.csv("ewidata.csv", row.names = 1)
+# load updated ewidata!
+ewidata <- read.csv("ewidata.csv", row.names=1)
 
-levels(dat$system)
+codes <- unique(ewidata$code)
 
-# restrict to GOA 
+# limit to GOA
+GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
+unique(GOA$code)
 
-goa.dat <- dat %>%
-  filter(system %in% c("EGOA", "WGOA"))
+# drop CI pinks as they aren't present in odd and even years
+GOA <- filter(GOA, code != "CI.PI")
 
-# drop climate data!
-drop <- grep("AKC", goa.dat$code)
-goa.dat <- goa.dat[-drop,]
+# drop climate
+drop <- grep("AKCLIM", GOA$code)
+GOA <- GOA[-drop,]
 
-ggplot(goa.dat, aes(year, value)) +
+unique(GOA$code)
+
+# drop dogfish (too many NAs)
+drop <- grep("DOGF", GOA$code)
+GOA <- GOA[-drop,]
+
+# drop birds and icy strait and chatnam and seward line and CPR
+drop <- c(grep("PROD.", GOA$code), grep("PHEN.", GOA$code), grep("ICY", GOA$code), grep("CHAT", GOA$code), grep("SEWARD", GOA$code), 
+          grep("mesozoo", GOA$code), grep("diatoms", GOA$code), grep("copepod", GOA$code))
+GOA <- GOA[-drop,]
+
+# herring time series are repeats, remove
+drop <- grep("SE.HER.ROE", GOA$code)
+GOA <- GOA[-drop,]
+
+# and drop the ichthyo data
+drop <- grep("ICH", GOA$code)
+GOA <- GOA[-drop,]
+
+# qnd, finally, drop PWS pinks as they are so hatchery-dominated
+GOA <- GOA %>%
+  filter(code != "PWS.PI")
+
+# and start in 1972
+GOA <- GOA[GOA$year>=1972,]
+
+unique(GOA$code)
+
+ggplot(GOA, aes(year, value)) +
   theme_bw() +
   geom_line() +
   geom_point() +
   facet_wrap(~code, scales="free_y") 
 
-# examine salmon timeseries alone
+# update with latest salmon values
+update <- read.csv("salmon catch updates.csv")
[…]

### 2019-09-04 paper draft.Rmd
New plotting for GOA biology

@@
 new.names <- c("East.spring.SST", "East.winter.SST", "SLP.gradient", "Papa.advection", "GAK1.salinity", "West.spring.SST", "West.winter.SST", "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149")
-
-pdf("goa.climate.loadings.pdf", 6,6)
-print(plot_loadings(rotated, names = new.names, threshold=0.8))
-dev.off()
 
@@
 # trends
-# pdf("goa.climate.trends.pdf", 6,6)
-# print(plot_trends(rotated, years = as.numeric(colnames(Y))))
-# dev.off()
+plot.trend <- data.frame(year=1972:2017, trend=as.vector(rotated$trends_mean),
+                         lo=as.vector(rotated$trends_lower),
+                         hi=as.vector(rotated$trends_upper))
+
+plot.b <- ggplot(plot.trend, aes(year, trend)) +
+  theme_bw() +
+  geom_line(color=cb[3]) +
+  geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
+  geom_hline(yintercept = 0) + ylab("Trend value") +
+  theme(axis.title.x = element_blank()) + 
+  ggtitle("b) Trend")
+
+png("combined GOA climate loading and trend plot.png", 8, 4, units = "in", res=300)
+
+ggpubr::ggarrange(plot.a, plot.b, ncol=2)
+
+dev.off()
 ```
@@
 
-# loadings
-pdf("goa.biol.loadings.pdf", 6, 6)
-print(plot_loadings(rotated, names = new.names, threshold=0.9))
-dev.off()
+# back to the bespoke code!
+loadings.1 <- as.data.frame(rotated$Z_rot[,,1])
+names(loadings.1)  <- new.names
+
+# drop columns that don't meet threshold
+# first, figure out the proportion above/below 0
+f <- function(x) max(sum(x<0)/length(x), sum(x>0)/length(x))
+
+prop1 <- apply(loadings.1, 2, f)
+prop1
+keep <- prop1 > 0.95
+
+loadings.1 <- loadings.1[,keep]
+
+# plot
+plot.load <- loadings.1 %>%
+  gather()
+
+rank <- tapply(plot.load$value, plot.load$key, mean)
+plot.load$rank <- rank[match(plot.load$key, names(rank))]
+
+plot.load$key <- reorder(plot.load$key, plot.load$rank)
+
+# set pallette
+cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+plot.a <- ggplot(plot.load, aes(key, value)) +
+  theme_bw() +
+  geom_violin(fill=cb[3]) +
+  geom_hline(yintercept = 0) +
+  theme(axis.title.y = element_blank()) + ylab("Loading") +
+  coord_flip() +
+  ggtitle("a) Loadings")
 ```
@@
 # trends
-
-pdf("goa.biol.trend.pdf", 6,6)
-print(plot_trends(rotated, years = as.numeric(colnames(Y))))
-dev.off()
+plot.trend <- data.frame(year=1972:2017, trend=as.vector(rotated$trends_mean),
+                         lo=as.vector(rotated$trends_lower),
+                         hi=as.vector(rotated$trends_upper))
+
+plot.b <- ggplot(plot.trend, aes(year, trend)) +
+  theme_bw() +
+  geom_line(color=cb[3]) +
+  geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
+  geom_hline(yintercept = 0) + ylab("Trend value") +
+  theme(axis.title.x = element_blank()) + 
+  ggtitle("b) Trend")
+
+png("combined GOA climate loading and trend plot.png", 8, 4, units = "in", res=300)
+
+ggpubr::ggarrange(plot.a, plot.b, ncol=2)
+
+dev.off()
+```
+
+```{r, echo=F}
+# look at rolling correlations between fitted and observed values, as in Ecology paper
+modelfit <- GOA.biol$best_model
+
+  n_ts <- dim(modelfit$data)[1]
+  n_years <- dim(modelfit$data)[2]
+  pred <- predicted(modelfit)
+   res.df <- data.frame(ID = rep(seq_len(n_ts),
+                                 n_years),
+                        Time = sort(rep(seq_len(n_years), n_ts)), 
+                        mean = c(t(apply(pred, c(3, 4), mean))), 
+                        lo = c(t(apply(pred, c(3, 4), quantile, 0.025))),
+                        hi = c(t(apply(pred, c(3, 4), quantile, 0.975))), 
+                        y = c(modelfit$data))
+
+        res.df$ID <- new.names
+
+  res.df$year <-  rep(1972:2017, each=n_ts)
+
+  res.df$mean.res <- res.df$y - res.df$mean
+  res.df$lo.res <- res.df$y - res.df$lo
+  res.df$hi.res <- res.df$y - res.df$hi
+
+ plot.these <- names(keep)[keep]
+
+ res.df <- res.df %>%
+   filter(ID %in% plot.these)
+
+  ggplot(res.df, aes_string(x = "year", y = "mean.res")) + geom_ribbon(aes_string(ymin = "lo.res", 
+        ymax = "hi.res"), fill=cb[3], alpha = 0.4) + geom_line(color=cb[3]) + facet_wrap("ID", 
+        scales = "free_y") + xlab("Year") + ylab("") + geom_hline(yintercept = 0, col=cb[2]) +
+        geom_vline(xintercept = c(1988.5, 2014), lty=2)
+
+# acf plot
+    index <- unique(res.df$ID)
+
+    par(mfrow=c(2,4))
+    for(i in 1:length(index)){
+     # i <- 1
+      temp <- res.df %>%
+        filter(ID==index[i])
+      temp <- na.omit(temp)
+      acf(temp$mean.res, main=index[i]) 
+    }
+# using 25-yr rolling windows!
+
+    dfa.cor <- data.frame()
+
+    for(i in 1:length(index)){    
+      # i <- 1 
+      temp <- res.df %>%
+        filter(ID==index[i])
+
+       for(ii in 13:58){
+       #  ii <- 13
+          win <- temp[(ii-12):(ii+12),]
+          temp.cor <- data.frame(ID=unique(win$ID),
+                                 year=win$year[13], 
+                                 window=25,
+                                 mean.cor=cor(win$y, win$mean, use="p"),
+                                 n.cor=sum(!is.na(win$y)))
+          dfa.cor <- rbind(dfa.cor, temp.cor)
+       }  }
+
+# now the 11-yr windows
+     for(i in 1:length(index)){    
+      # i <- 1 
+      temp <- res.df %>%
+        filter(ID==index[i])
+             for(ii in 6:65){
+       #  ii <- 13
+          win <- temp[(ii-5):(ii+5),]
+          temp.cor <- data.frame(ID=unique(win$ID),
+                                 year=win$year[6],
+                                 window=11,
+                                 mean.cor=cor(win$y, win$mean, use="p"),
+                                 n.cor=sum(!is.na(win$y)))
+          dfa.cor <- rbind(dfa.cor, temp.cor)
+             }  }
+
+      # and the 15-yr windows
+    for(i in 1:length(index)){    
+      # i <- 1 
+      temp <- res.df %>%
+        filter(ID==index[i])
+
+             for(ii in 8:63){
+       #  ii <- 13
+          win <- temp[(ii-7):(ii+7),]
+          temp.cor <- data.frame(ID=unique(win$ID),
+                                 year=win$year[8],
+                                 window=15,
+                                 mean.cor=cor(win$y, win$mean, use="p"),
+                                 n.cor=sum(!is.na(win$y)))
+          dfa.cor <- rbind(dfa.cor, temp.cor)
+       }  }
+
+# remove correlations with < 10 obs
+dfa.cor <- dfa.cor %>%
+  filter(n.cor>=10)
+ dfa.cor$window <- as.factor(dfa.cor$window)
+
+dfa.cor <- na.omit(dfa.cor)
+
+ggplot(dfa.cor, aes(year, mean.cor, color=window)) + 
+  theme_bw() + theme(axis.title.x = element_blank()) +
+  geom_line() + facet_wrap(~ID, 
+        scales = "free_y") + xlab("Year") + ylab("Correlation")
+
+ggsave("goa biology dfa correlations 11 15 25 year windows.png", height = 4, width=6, units="in")
+
+# and plot 15-yr windows
+ggplot(filter(dfa.cor, window==15), aes(year, mean.cor)) + 
+  theme_bw() + theme(axis.title.x = element_blank()) +
+  geom_line(color=cb[3]) + facet_wrap(~ID, 
+        scales = "free_y") + xlab("Year") + ylab("Correlation")
+
+ggsave("goa biology dfa correlations 15 year windows.png", height = 4, width=6, units="in")
+
+# and plot 25-yr windows
+ggplot(filter(dfa.cor, window==25), aes(year, mean.cor)) + 
+  theme_bw() + theme(axis.title.x = element_blank()) +
+  geom_line(color=cb[3]) + facet_wrap(~ID, 
+        scales = "free_y") + xlab("Year") + ylab("Correlation")
+
+ggsave("goa biology dfa correlations 25 year windows.png", height = 5.5, width=8, units="in")
+
 ```

### 2019-09-05 paper draft.Rmd
Tweaks to biol plots

@@
 library(rstanarm)
-
+library(tidyverse)
 ```
@@
   theme_bw() + theme(axis.title.x = element_blank()) +
-  geom_line(color=cb[3]) + facet_wrap(~ID, 
+  geom_path(color=cb[3]) + geom_point(color=cb[3]) + facet_wrap(~ID, 
         scales = "free_y") + xlab("Year") + ylab("Correlation")
@@
 
-```{r, include=F}
+```{r}
 

### 2019-09-09 paper draft.Rmd
Office updates

@@
 ```{r, echo=F}
-GOA.biol <- readRDS("GOA_biol_9.4.19.rds") 
+GOA.biol <- readRDS("GOA_biol_9.9.19.rds") 
 GOA.biol$summary
@@
 
-new.names <- c("Chignik.coho", "Chignik.pink", "Chiniak.eulachon", "Chiniak.jellyfish", "Chiniak.shrimp", "Cook.Inlet.coho", "Kodiak.coho", "Kodiak.pink", "Pavlof.capelin", "Pavlof.eulachon", "Pavlof.jellyfish", "Pavlof.shrimp", "Prince.William.coho", "Southeast.coho", "Southeast.herring", "Southeast.pink", "South.Peninsula.coho", "South.Peninsula.pink")
+new.names <- c("Chignik.coho", "Chignik.pink", "Chiniak.eulachon", "Chiniak.jellyfish", "Chiniak.P.cod", "Chiniak.shrimp", "Cook.Inlet.coho", "Kodiak.coho", "Kodiak.pink", "Pavlof.capelin", "Pavlof.eulachon", "Pavlof.jellyfish", "Pavlof.P.cod", "Pavlof.shrimp", "Prince.William.coho", "Southeast.coho", "Southeast.herring", "Southeast.pink", "South.Peninsula.coho", "South.Peninsula.pink")
+
+ check <- data.frame(names=names, newnames=new.names)
+ check
 
@@
 
-ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
-keep <- grep("AKCLIM_GOA", ewidata$code)
-clim.var <- ewidata[keep,]  
-keep <- grep("win.sst", clim.var$code)
-clim.var <- clim.var[keep,]
+ewidata <- read.csv("updated goa climate data.csv", row.names=1)
+
+keep <- grep("win.sst", colnames(ewidata))
+
+clim.var <- ewidata[,keep] %>%
+  mutate(year=1950:2019) %>%
+  gather(key, value, -year)
 
@@
 cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
-cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
+cli_dat$slope[cli_dat$Year %in% 2014:2017] <- 3
 cli_dat$slope <- as.factor(cli_dat$slope)
@@
 set.seed(99)
-mod = trend_lm(rotated_modelfit = r, all_years = 1972:2016,
-  selected_years = 1972:2016, climate_dat=cli_dat)
-
-names(mod)[2:4] <- c("1972-1988", "1989-2013", "2014-2016")
+mod = trend_lm(rotated_modelfit = r, all_years = 1972:2017,
+  selected_years = 1972:2017, climate_dat=cli_dat)
+
+names(mod)[2:4] <- c("1972-1988", "1989-2013", "2014-2017")
 summary(mod)
@@
 
-melt.mod <- melt(mod, measure.vars=c("1972-1988", "1989-2013", "2014-2016"))
+melt.mod <- melt(mod, measure.vars=c("1972-1988", "1989-2013", "2014-2017"))
 names(melt.mod)[3] <- "era"
-
-pdf("Fig. 3A goa.biol.on.sst.pdf", 6, 4)
-ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="A) SST")
-dev.off()
+melt.mod$covar <- "SST"
 
@@
 
-years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2016
+years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2017
 
 # set up climate data
-cli_dat <- data.frame(Year=pdo$YEAR,PDO=win.pdo)
+cli_dat <- data.frame(Year=pdo$YEAR, PDO=win.pdo)
 
@@
 set.seed(99)
-mod2 = trend_lm(rotated_modelfit = r, all_years = 1972:2016,
+mod2 = trend_lm(rotated_modelfit = r, all_years = 1972:2017,
   selected_years = 1972:2016, climate_dat=cli_dat)
 
-names(mod2)[2:4] <- c("1972-1988", "1989-2013", "2014-2016")
+names(mod2)[2:4] <- c("1972-1988", "1989-2013", "2014-2017")
 
@@
 
-melt.mod <- melt(mod2, measure.vars=c("1972-1988", "1989-2013", "2014-2016"))
-names(melt.mod)[3] <- "era"
-
-pdf("Fig. 3B goa.biol.on.pdo.pdf", 6, 4)
-ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) PDO")
-dev.off()
-
-```
-
-And the NPGO.
-
-```{r, echo=F}
-
-# 
+melt.mod2 <- melt(mod2, measure.vars=c("1972-1988", "1989-2013", "2014-2017"))
+names(melt.mod2)[3] <- "era"
+melt.mod2$covar <- "PDO"
+
+```
+
+And the NPGO. Leaving this commented out as there is no meaningfull effect of the NPGO in any era!
+
+```{r, echo=F}
+
 # npgo <- read.csv("winter.npgo.csv")
 # 
-# years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2016
+# years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2017
 # 
@@
 # cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
-# cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
+# cli_dat$slope[cli_dat$Year %in% 2014:2017] <- 3
 # cli_dat$slope <- as.factor(cli_dat$slope)
@@
 # 
-# mod2 = trend_lm(rotated_modelfit = r, all_years = 1972:2016,
-#   selected_years = 1972:2016, climate_dat=cli_dat)
+# mod3 = trend_lm(rotated_modelfit = r, all_years = 1972:2017,
+#   selected_years = 1972:2017, climate_dat=cli_dat)
 # 
 # 
-# summary(mod2)
+# summary(mod3)
 # 
-# names(mod2)[2:4] <- c("1972-1988", "1989-2013", "2014-2016")
+# names(mod3)[2:4] <- c("1972-1988", "1989-2013", "2014-2017")
 # 
-# melt.mod <- melt(mod2, measure.vars=c("1972-1988", "1989-2013", "2014-2016"))
-# names(melt.mod)[3] <- "era"
+# melt.mod3 <- melt(mod3, measure.vars=c("1972-1988", "1989-2013", "2014-2017"))
+# names(melt.mod3)[3] <- "era"
+# melt.mod3$covar <- "NPGO"
 # 
 # pdf("Fig. 3C goa.biol.on.npgo.pdf", 6, 4)
-# ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) NPGO")
+# ggplot(melt.mod3, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="C) NPGO")
 # dev.off()
 
+```
+
+Combine and plot
+```{r}
+plot.melt <- rbind(melt.mod, melt.mod2)
+plot.melt$covar <- reorder(plot.melt$covar, desc(plot.melt$covar))
+plot.melt$era <- reorder(plot.melt$era, desc(plot.melt$era))
+ggplot(plot.melt, aes(x=era, y=value)) +
+  theme_bw() + 
+  geom_violin(fill=cb[3]) +
+  geom_hline(yintercept=0, lty=2) +
+  geom_vline(xintercept = 0, lty=2) +
+  facet_wrap(~covar) +
+  theme(axis.title.y = element_blank()) +
+  ylab("Slope") +
+  coord_flip()
+
+ggsave("goa biology era sst pdo slopes.png", width=6, height=4, units="in")
 ```
@@
 rotated = rotate_trends(GOA.ich$best_model) 
-
 new.names <- c("sandlance", "ronquils", "walleye.pollock", "Pacific.cod", "flathead.sole", "n.rock.sole", "rockfish")
 
-# loadings
-pdf("goa.ich.loadings.pdf", 6, 6)
-print(plot_loadings(rotated, names = new.names, threshold=0.9))
-dev.off()
+ check <- data.frame(names=names, newnames=new.names)
+ check
+
+ # back to the bespoke code!
+loadings.1 <- as.data.frame(rotated$Z_rot[,,1])
+names(loadings.1)  <- new.names
+
+# drop columns that don't meet threshold
+# first, figure out the proportion above/below 0
+f <- function(x) max(sum(x<0)/length(x), sum(x>0)/length(x))
+
+prop1 <- apply(loadings.1, 2, f)
+prop1
+keep <- prop1 > 0.95
+
+loadings.1 <- loadings.1[,keep]
+
+# plot
+plot.load <- loadings.1 %>%
+  gather()
+
+rank <- tapply(plot.load$value, plot.load$key, mean)
+plot.load$rank <- rank[match(plot.load$key, names(rank))]
+
+plot.load$key <- reorder(plot.load$key, plot.load$rank)
+
+# set pallette
+cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+plot.a <- ggplot(plot.load, aes(key, value)) +
+  theme_bw() +
+  geom_violin(fill=cb[3]) +
+  geom_hline(yintercept = 0) +
+  theme(axis.title.y = element_blank()) + ylab("Loading") +
+  coord_flip() +
+  ggtitle("a) Ichthyoplankton loadings")
+
+plot.a
+
 ```
@@
 # trends
-
-pdf("goa.ich.trend.pdf", 6,6)
-print(plot_trends(rotated, years = as.numeric(colnames(Y))))
-dev.off()
+plot.trend <- data.frame(year=as.numeric(as.character(colnames(Y))), trend=as.vector(rotated$trends_mean),
+                         lo=as.vector(rotated$trends_lower),
+                         hi=as.vector(rotated$trends_upper))
+
+plot.b <- ggplot(plot.trend, aes(year, trend)) +
+  theme_bw() +
+  geom_line(color=cb[3]) +
+  geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
+  geom_hline(yintercept = 0) + ylab("Trend value") +
+  theme(axis.title.x = element_blank()) + 
+  ggtitle("b) Ichthyoplankton trend")
+
+plot.b
+# png("combined GOA climate loading and trend plot.png", 8, 4, units = "in", res=300)
+# 
+# ggpubr::ggarrange(plot.a, plot.b, ncol=2)
+# 
+# dev.off()
 ```
@@
 ```{r, include=F}
-
 #  load sst as the explanatory variable.
 
-ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
-keep <- grep("AKCLIM_GOA", ewidata$code)
-clim.var <- ewidata[keep,]  
-keep <- grep("win.sst", clim.var$code)
-clim.var <- clim.var[keep,]
+ewidata <- read.csv("updated goa climate data.csv", row.names=1)
+
+keep <- grep("win.sst", colnames(ewidata))
+
+clim.var <- ewidata[,keep] %>%
+  mutate(year=1950:2019) %>%
+  gather(key, value, -year)
 
@@
 
-pdf("Fig. x goa.ich.on.sst.pdf", 4, 5)
 ggplot(mod, aes(x=SST)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + xlab("Slope")
-dev.off()
 
@@
 
-# loadings
-pdf("goa.zoop.loadings.pdf", 6, 6)
-print(plot_loadings(rotated, names = new.names, threshold=0.9))
-dev.off()
+ # back to the bespoke code!
+loadings.1 <- as.data.frame(rotated$Z_rot[,,1])
+names(loadings.1)  <- new.names
+
+# drop columns that don't meet threshold
+# first, figure out the proportion above/below 0
+f <- function(x) max(sum(x<0)/length(x), sum(x>0)/length(x))
+
+prop1 <- apply(loadings.1, 2, f)
+prop1
+keep <- prop1 > 0.95
+
+loadings.1 <- loadings.1[,keep]
+
+# plot
+plot.load <- loadings.1 %>%
+  gather()
+
+rank <- tapply(plot.load$value, plot.load$key, mean)
+plot.load$rank <- rank[match(plot.load$key, names(rank))]
+
+plot.load$key <- reorder(plot.load$key, plot.load$rank)
+
+# set pallette
+cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+plot.d <- ggplot(plot.load, aes(key, value)) +
+  theme_bw() +
+  geom_violin(fill=cb[3]) +
+  geom_hline(yintercept = 0) +
+  theme(axis.title.y = element_blank()) + ylab("Loading") +
+  coord_flip() +
+  ggtitle("d) Zooplankton loadings")
+
+plot.d
 ```
@@
 # trends
-
-pdf("goa.zoop.trend.pdf", 6,6)
-print(plot_trends(rotated, years = as.numeric(colnames(Y))))
-dev.off()
+plot.trend <- data.frame(year=as.numeric(as.character(colnames(Y))), trend=as.vector(rotated$trends_mean),
+                         lo=as.vector(rotated$trends_lower),
+                         hi=as.vector(rotated$trends_upper))
+
+plot.e <- ggplot(plot.trend, aes(year, trend)) +
+  theme_bw() +
+  geom_line(color=cb[3]) +
+  geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
+  geom_hline(yintercept = 0) + ylab("Trend value") +
+  theme(axis.title.x = element_blank()) + 
+  ggtitle("e) Zooplankton trend")
+
+plot.e
 ```
@@
 ```{r, include=F}
-
 #  load sst as the explanatory variable.
-
-ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
-keep <- grep("AKCLIM_GOA", ewidata$code)
-clim.var <- ewidata[keep,]  
-keep <- grep("win.sst", clim.var$code)
-clim.var <- clim.var[keep,]
+ewidata <- read.csv("updated goa climate data.csv", row.names=1)
+
+keep <- grep("win.sst", colnames(ewidata))
+
+clim.var <- ewidata[,keep] %>%
+  mutate(year=1950:2019) %>%
+  gather(key, value, -year)
 
@@
 set.seed(99)
-mod = trend_lm(rotated_modelfit = r, all_years = years,
+mod2 = trend_lm(rotated_modelfit = r, all_years = years,
   selected_years = years, climate_dat=cli_dat)
 
-summary(mod)
+summary(mod2)
 
 # get 95% CIs
-x <- sort(mod[,2])
-print(names(mod)[2])
+x <- sort(mod2[,2])
+print(names(mod2)[2])
 print(paste("median: ", median(x)))
@@
 
-pdf("Fig. x goa.zoop.on.sst.pdf", 4, 5)
-ggplot(mod, aes(x=SST)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + xlab("Slope")
-dev.off()
+mod2$covar <- "SST"
+
+ggplot(mod2, aes(x=covar, y=SST)) + 
+  theme_bw() +
+  geom_violin() + 
+  geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + xlab("Slope")
 

### 2019-09-09 ewidata update - GOA biology.R

@@ -5,6 +5,7 @@ library(fields)
 library(maps)
 library(mapdata)
 library(zoo)
+library(gtools)
 
 # load updated ewidata!
 ewidata <- read.csv("ewidata.csv", row.names=1)
@@ -123,6 +124,18 @@ ggplot(scaled.pink, aes(year, value)) +
 # put back into GOA
 GOA <- rbind(GOA, scaled.pink)
 
+# load small-mesh P. cod cpue
+
+chini <- read.csv("chiniak.sm.mesh.csv", row.names = 1)
+
+xtra1 <- data.frame(year=chini$year, code="CHI.P.COD.CPUE", value=chini$p.cod, system=NA, subtype=NA)
+
+pav <- read.csv("pav.small mesh cpue.csv", row.names = 1)
+
+xtra2 <- data.frame(year=pav$year, code="PAV.P.COD.CPUE", value=pav$p.cod, system=NA, subtype=NA)
+
+GOA <- rbind(GOA, xtra1, xtra2)
+
 # and save
 write.csv(GOA, "updated GOA biology data.csv", row.names = F)
 
@@ -155,7 +168,7 @@ options(mc.cores = parallel::detectCores())
 sub_data = GOA
 
 # set the name for the model output!
-name <- "GOA_biol_9.4.19"
+name <- "GOA_biol_9.9.19"
 
 # reshape data
 melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))

### 2019-09-12 paper draft.Rmd
Changes to plots

@@
 ```{r, echo=F}
-GOA.ich <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_ich_7.19.18.one.trend.rds") 
+GOA.ich <- readRDS("/Users/MikeLitzow 1/Documents/R/Fate/fate-ewi/GOA_ich_7.19.18.one.trend.rds") 
 GOA.ich$summary
@@
   coord_flip() +
-  ggtitle("a) Ichthyoplankton loadings")
+  ggtitle("a) Ichthyoplankton loadings") +
+  theme(title = element_text(size=8))
 
@@
   geom_hline(yintercept = 0) + ylab("Trend value") +
-  theme(axis.title.x = element_blank()) + 
-  ggtitle("b) Ichthyoplankton trend")
+  xlab("Year") +
+  ggtitle("b) Ichthyoplankton trend") +
+  theme(title = element_text(size=8))
 
@@
 
-ggplot(mod, aes(x=SST)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + xlab("Slope")
+plot.c <- ggplot(mod, aes(x=SST)) +
+  theme_bw() +
+  geom_density(fill=cb[3]) + 
+  geom_hline(yintercept=0) + 
+  geom_vline(xintercept = 0, lty=2) + 
+  xlab("Slope") + 
+  ylab("Density") +
+  ggtitle(("c) Ichthyoplankton slope on SST")) +
+  theme(title=element_text(size=8))
 
@@
 ```{r, echo=F}
-GOA.zoop <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_zoop_7.19.18.one.trend.rds") 
+GOA.zoop <- readRDS("/Users/MikeLitzow 1/Documents/R/Fate/fate-ewi/GOA_zoop_7.19.18.one.trend.rds") 
 GOA.zoop$summary
@@
   coord_flip() +
-  ggtitle("d) Zooplankton loadings")
+  ggtitle("d) Zooplankton loadings") +
+  theme(title = element_text(size=8))
 
@@
   geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
-  geom_hline(yintercept = 0) + ylab("Trend value") +
-  theme(axis.title.x = element_blank()) + 
-  ggtitle("e) Zooplankton trend")
+  geom_hline(yintercept = 0) + ylab("Trend value") + xlab("Year") +
+  ggtitle("e) Zooplankton trend") +
+  theme(title = element_text(size=8))
 
@@
 set.seed(99)
-mod2 = trend_lm(rotated_modelfit = r, all_years = years,
+mod = trend_lm(rotated_modelfit = r, all_years = years,
   selected_years = years, climate_dat=cli_dat)
 
-summary(mod2)
+summary(mod)
 
 # get 95% CIs
-x <- sort(mod2[,2])
-print(names(mod2)[2])
+x <- sort(mod[,2])
+print(names(mod)[2])
 print(paste("median: ", median(x)))
@@
 
-mod2$covar <- "SST"
-
-ggplot(mod2, aes(x=covar, y=SST)) + 
+plot.f <- ggplot(mod, aes(x=SST)) +
   theme_bw() +
-  geom_violin() + 
-  geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + xlab("Slope")
-
-```
-
-# Now the EBS results.
-
-```{r, echo=F}
-
-EBS.clim <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/EBS_clim_6.19.18.rds") # read in EBS climate model
-
-EBS.clim$summary
-```
-
-And the trends and loading plots.
-
-```{r, echo=F}
-
-# recover the years and names for plots!
-ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
-keep <- grep("AKCLIM_EBS", ewidata$code)
-
-ewidata <- ewidata[keep,]
-sub_data = ewidata
-
-melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
-Y <- dcast(melted, code ~ year)
-names = Y$code
-Y = as.matrix(Y[,-which(names(Y) == "code")])
-
-rotated = rotate_trends(EBS.clim$best_model)
-
-new.names <- c("Bottom.temp", "Spring.SST", "Winter.SST", "Summer.NW.wind", "Summer.SE.wind", "Winter.NW.wind", "Winter.SE.wind", "Ice.57.75N", "Ice.58.75N", "Ice.64.25N")
-
-# loadings
-pdf("ebs.clim.loadings.pdf", 6, 6)
-print(plot_loadings(rotated, names = new.names, threshold=0.9))
+  geom_density(fill=cb[3]) + 
+  geom_hline(yintercept=0) + 
+  geom_vline(xintercept = 0, lty=2) + 
+  xlab("Slope") + 
+  ylab("Density") +
+  ggtitle(("f) Zooplankton slope on SST")) +
+  theme(title = element_text(size=8))
+
+```
+
+And combine into a single plot.
+
+```{r}
+png("combined short time series plots.png", 8,5, units="in", res=300)
+
+ggpubr::ggarrange(plot.a, plot.b, plot.c, plot.d, plot.e, plot.f,
+          ncol=3, widths=c(1, 1, 0.8), nrow=2)
+
 dev.off()
-
-# trends
-pdf("ebs.clim.trend.pdf", 6,6)
-print(plot_trends(rotated, years = as.numeric(colnames(Y))))
-dev.off()
-```
-
-Find swans.
-
-```{r, echo=F}
-xx <- find_swans(rotated, threshold = 0.01, plot=F)
-xx$year <- 1950:2016
-xx <- filter(xx, below_threshold == T)
-xx
-```
-
-Test for regime probability. Use n=3 maximum states to rule out a switch to a warmer state.
-
-```{r, include=F}
-
-y = apply(rotated$trends, c(2,3), mean)
-sd_y = apply(rotated$trends, c(2,3), sd)
-
-# run each trend sequentially through the regimes
-# code to identify (1) number of changes and (2)
-# where those change points occur.
-set.seed(99)
-f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)
-
-```
-
-Regime results for Trend 1. The 2-regime model is best-supported.
-
-```{r, echo=F}
-print(f1$table) # this shows 2-regime model is best
-plot_regime_model(f1$best_model)
-```
-
-## EBS biology results.
-
-```{r, echo=F}
-EBS.biol <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/EBS_biol_6.19.18.rds") 
-EBS.biol$summary
-```
-
-Now plot trends and loadings.
-
-```{r, echo=F}
-
-# reload the version of data used
-# load updated ewidata!
-ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
-
-# first, EBS biology
-codes <- unique(ewidata$code)
-
-# limit to EBS
-EBS <- ewidata[ewidata$system == "EBS",]
-
-# and start in 1975, when data beyond salmon are first available
-EBS <- EBS[EBS$year>=1975,]
-
-unique(EBS$code)
-
-# drop clim
-drop <- grep("AKCLIM", EBS$code)
-EBS <- EBS[-drop,]
-
-# drop depth
-drop <- grep("DEPTH", EBS$code)
-EBS <- EBS[-drop,]
-
-# recover the years and names for plots!
-sub_data = EBS
-
-melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
-Y <- dcast(melted, code ~ year)
-names = Y$code
-Y = as.matrix(Y[,-which(names(Y) == "code")])
-
-rotated = rotate_trends(EBS.biol$best_model)
-
-new.names <- c("Alaska.place.lat", "Bristol.Bay.coho", "Bering.flounder.lat", "St.George.kitt.phen", "St.Paul.kitt.phen", "St.Paul.kitt.product", "com.eider.eggs/nest", "com.eider.hatch.date", "com.eider.success", "St.George.com.murre.phenol", "St.Paul.com.murre.phenol", "St.George.com.murre.product", "St.Paul.com.murre.product", "juv.cod.cpue", "juv.pollock.cpue", "flathead.sole.lat", "gl.gull.eggs/nest", "gl.gull.hatch.date", "gl.gull.success", "Grnlnd.turbot.lat", "Pac.cod.lat", "Pac.halibut.lat", "red.king.crab.lat", "rock.sole.lat", "shortf.eelpout.lat", "snow.crab.lat", "spec.eider.eggs/nest", "spec.eider.hatch.date", "spec.eider.success", "sturgeon.poacher.lat", "Tanner.crab.lat", "St.George.tb.murre.phenol", "St.Paul.tb.murre.phenol", "St.George.tb.murre.product", "St.Paul.tb.murre.product", "walleye.pollock.lat", "wattled.eelpout.lat", "yellowfin.sole.lat")
-
-# loadings
-pdf("ebs.biol.loadings.pdf", 6, 6)
-print(plot_loadings(rotated, names = new.names, threshold=0.9))
-dev.off()
-
-```
-
-```{r, echo=F}
-# trends
-
-pdf("ebs.biol.trend.pdf", 6,6)
-print(plot_trends(rotated, years = as.numeric(colnames(Y))))
-dev.off()
-```
-Test for regime probability. Use n=3 maximum states to rule out a switch to a warmer state.
-
-```{r, include=F}
-
-y = apply(rotated$trends, c(2,3), mean)
-sd_y = apply(rotated$trends, c(2,3), sd)
-
-# run each trend sequentially through the regimes
-# code to identify (1) number of changes and (2)
-# where those change points occur.
-set.seed(99)
-f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)
-
-```
-
-Regime results for Trend 1. The 2-regime model is best-supported.
-
-```{r, echo=F}
-print(f1$table) # this shows 2-regime model is best
-plot_regime_model(f1$best_model)
-```
-
-Next step is to examine relationships between SST and the EBS biology trend.
-
-```{r, message=FALSE, results='hide', echo=F}
-
-# load sst as the explanatory variable.
-
-ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
-keep <- grep("AKCLIM_EBS", ewidata$code)
-clim.var <- ewidata[keep,]  
-keep <- grep("win.sst", clim.var$code)
-clim.var <- clim.var[keep,]
-
-win.sst <- tapply(clim.var$value, clim.var$year, identity)
-
-years <- as.numeric(colnames(EBS.biol$best_model$data)) # these are the years for the biology model - 1975:2016
-
-cli_dat <- data.frame(Year=as.numeric(names(win.sst)), SST=win.sst)
-
-cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))
-
-# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
-
-cli_dat$slope <- 1
-cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
-cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
-cli_dat$slope <- as.factor(cli_dat$slope)
-
-# rename rotated
-r <- rotated
-
-Y = rnorm(length(cli_dat$slope))
-cli_dat = as.data.frame(model.matrix(Y ~ Year + SST:slope, data=cli_dat))
-# replace ':' in name for formula
-names(cli_dat) = gsub(":","",names(cli_dat))
-# remove named intercept
-cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
-
-set.seed(4)
-mod = trend_lm(rotated_modelfit = r, all_years = 1975:2016,
-  selected_years = 1975:2016, climate_dat=cli_dat)
-
-names(mod)[2:4] <- c("1975-1988", "1989-2013", "2014-2016")
-summary(mod)
-
-# get 95% CIs
-
-for(i in 2:4){
-x <- sort(mod[,i])
-print(names(mod)[i])
-print(paste("median: ", median(x)))
-print(paste("LCI: ", x[(0.025*25000)]))
-print(paste("UCI: ", x[(0.975*25000)]))
-}
-
-melt.mod <- melt(mod, measure.vars=c("1975-1988", "1989-2013", "2014-2016"))
-names(melt.mod)[3] <- "era"
-
-pdf("Fig. x ebs.biol.on.sst.pdf", 6, 4)
-ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="A) SST")
-dev.off()
-```
-
-Now the PDO.
-
-```{r, message=FALSE, results='hide', echo=F}
-
-pdo <- read.csv("pdo.csv")
-
-win.pdo <- pdo$NDJFM
-
-years <- as.numeric(colnames(EBS.biol$best_model$data)) # these are the years for the biology model - 1972:2016
-
-# set up climate data
-cli_dat <- data.frame(Year=pdo$YEAR,PDO=win.pdo)
-
-cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(PDO = arm::rescale(PDO))
-
-# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
-
-cli_dat$slope <- 1
-cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
-cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
-cli_dat$slope <- as.factor(cli_dat$slope)
-
-# rename rotated
-r <- rotated
-
-Y = rnorm(length(cli_dat$slope))
-cli_dat = as.data.frame(model.matrix(Y ~ Year + PDO:slope, data=cli_dat))
-# replace ':' in name for formula
-names(cli_dat) = gsub(":","",names(cli_dat))
-# remove named intercept
-cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
-
-set.seed(99)
-mod2 = trend_lm(rotated_modelfit = r, all_years = 1975:2016,
-  selected_years = 1975:2016, climate_dat=cli_dat)
-
-names(mod2)[2:4] <- c("1975-1988", "1989-2013", "2014-2016")
-
-summary(mod2)
-
-# get 95% CIs
-
-for(i in 2:4){
-x <- sort(mod2[,i])
-print(names(mod2)[i])
-print(paste("median: ", median(x)))
-print(paste("LCI: ", x[(0.025*25000)]))
-print(paste("UCI: ", x[(0.975*25000)]))
-}
-
-melt.mod <- melt(mod2, measure.vars=c("1975-1988", "1989-2013", "2014-2016"))
-names(melt.mod)[3] <- "era"
-
-pdf("Fig. x ebs.biol.on.pdo.pdf", 6, 4)
-ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) PDO")
-dev.off()
-
-```
-
-And the winter AO.
-
-```{r, echo=F}
-
-ao <- read.csv("winter.ao.csv")
-
-win.ao <- ao$NDJFM
-
-years <- as.numeric(colnames(EBS.biol$best_model$data)) # these are the years for the biology model - 1975:2016
-
-# set up climate data
-cli_dat <- data.frame(Year=ao$YEAR,AO=win.ao)
-
-cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(AO = arm::rescale(AO))
-
-# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
-
-cli_dat$slope <- 1
-cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
-cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
-cli_dat$slope <- as.factor(cli_dat$slope)
-
-# rename rotated
-r <- rotated
-
-Y = rnorm(length(cli_dat$slope))
-cli_dat = as.data.frame(model.matrix(Y ~ Year + AO:slope, data=cli_dat))
-# replace ':' in name for formula
-names(cli_dat) = gsub(":","",names(cli_dat))
-# remove named intercept
-cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
-
-set.seed(99)
-mod3 = trend_lm(rotated_modelfit = r, all_years = 1975:2016,
-  selected_years = 1975:2016, climate_dat=cli_dat)
-
-names(mod3)[2:4] <- c("1975-1988", "1989-2013", "2014-2016")
-
-summary(mod3)
-
-# get 95% CIs
-
-for(i in 2:4){
-x <- sort(mod3[,i])
-print(names(mod3)[i])
-print(paste("median: ", median(x)))
-print(paste("LCI: ", x[(0.025*25000)]))
-print(paste("UCI: ", x[(0.975*25000)]))
-}
-
-# and proportion > 0
-for(i in 2:4){
-x <- sort(mod3[,i])
-print(names(mod3)[i])
-print(paste("prop > 0: ", sum(x > 0)/length(x)))
-}
-
-melt.mod <- melt(mod3, measure.vars=c("1975-1988", "1989-2013", "2014-2016"))
-names(melt.mod)[3] <- "era"
-
-pdf("Fig. x ebs.biol.on.ao.pdf", 6, 4)
-ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) AO")
-dev.off()
-```
-
-```{r}
-head(melt.mod)
-restricted <- filter(melt.mod, era %in% c("1975-1988", "1989-2013"))
-restricted <- melt.mod
-png("ebs community response to ao.png", 6,4, units="in", res=300)
-ggplot(data=restricted, aes(fill=era, x=era, y=value)) + geom_violin() + coord_flip() + ggtitle("Community response to Arctic Oscillation") +
-  geom_hline(yintercept = 0) + xlab("Era") + ylab("Regression coefficient") + theme_gray() + theme(legend.position = c(0.8, 0.5), legend.title = element_blank())
-dev.off()
-```
+```

### 2019-09-12 paper draft.Rmd
Tweaked short TS plot

@@
 ```{r, echo=F}
-GOA.ich <- readRDS("/Users/MikeLitzow 1/Documents/R/Fate/fate-ewi/GOA_ich_7.19.18.one.trend.rds") 
+GOA.ich <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_ich_7.19.18.one.trend.rds") 
 GOA.ich$summary
@@
   ylab("Density") +
-  ggtitle(("c) Ichthyoplankton slope on SST")) +
+  ggtitle(("c) Ichthyo. slope on SST")) +
   theme(title=element_text(size=8))
@@
 ```{r, echo=F}
-GOA.zoop <- readRDS("/Users/MikeLitzow 1/Documents/R/Fate/fate-ewi/GOA_zoop_7.19.18.one.trend.rds") 
+GOA.zoop <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_zoop_7.19.18.one.trend.rds") 
 GOA.zoop$summary

### 2019-09-17 paper draft.Rmd
Added rolling correlations for all climate time series
End of the day!

@@
 
+# re-run with al climate TS, not just the ones meeting the loadings threshold used above
+
+# set up res.df with all TS
+   res.df <- data.frame(ID = rep(seq_len(n_ts),
+                                 n_years),
+                        Time = sort(rep(seq_len(n_years), n_ts)), 
+                        mean = c(t(apply(pred, c(3, 4), mean))), 
+                        lo = c(t(apply(pred, c(3, 4), quantile, 0.025))),
+                        hi = c(t(apply(pred, c(3, 4), quantile, 0.975))), 
+                        y = c(modelfit$data))
+
+        res.df$ID <- new.names
+
+  res.df$year <-  rep(1950:2019, each=n_ts)
+
+  res.df$mean.res <- res.df$y - res.df$mean
+  res.df$lo.res <- res.df$y - res.df$lo
+  res.df$hi.res <- res.df$y - res.df$hi
+
+# look at rolling correlations between fitted and observed values, as in Ecology paper
+# using 25-yr rolling windows!
+ index <- unique(res.df$ID)
+    dfa.cor <- data.frame()
+
+    for(i in 1:length(index)){    
+      # i <- 1 
+      temp <- res.df %>%
+        filter(ID==index[i])
+
+       for(ii in 13:58){
+       #  ii <- 13
+          win <- temp[(ii-12):(ii+12),]
+          temp.cor <- data.frame(ID=unique(win$ID),
+                                 year=win$year[13], 
+                                 window=25,
+                                 mean.cor=cor(win$y, win$mean, use="p"),
+                                 n.cor=sum(!is.na(win$y)))
+          dfa.cor <- rbind(dfa.cor, temp.cor)
+       }  }
+
+# remove salinity correlations with < 10 obs
+dfa.cor <- dfa.cor %>%
+  filter(n.cor>=10)
+ dfa.cor$window <- as.factor(dfa.cor$window)
+
+ggplot(filter(dfa.cor, window==25), aes(year, mean.cor)) + 
+  theme_bw() + theme(axis.title.x = element_blank()) +
+  geom_line(color=cb[3]) + facet_wrap(~ID, 
+        scales = "free_y", ncol=3) + xlab("Year") + ylab("Correlation")
+
+ggsave("goa climate dfa correlations 25 year windows all TS.png", height = 5, width=6, units="in")    
 ```
@@
 
+# as for climate data, recalculate rolling correlations with all time series
+modelfit <- GOA.biol$best_model
+
+  n_ts <- dim(modelfit$data)[1]
+  n_years <- dim(modelfit$data)[2]
+  pred <- predicted(modelfit)
+   res.df <- data.frame(ID = rep(seq_len(n_ts),
+                                 n_years),
+                        Time = sort(rep(seq_len(n_years), n_ts)), 
+                        mean = c(t(apply(pred, c(3, 4), mean))), 
+                        lo = c(t(apply(pred, c(3, 4), quantile, 0.025))),
+                        hi = c(t(apply(pred, c(3, 4), quantile, 0.975))), 
+                        y = c(modelfit$data))
+
+        res.df$ID <- new.names
+
+  res.df$year <-  rep(1972:2017, each=n_ts)
+
+  res.df$mean.res <- res.df$y - res.df$mean
+  res.df$lo.res <- res.df$y - res.df$lo
+  res.df$hi.res <- res.df$y - res.df$hi
+
+modelfit <- GOA.biol$best_model
+
+  n_ts <- dim(modelfit$data)[1]
+  n_years <- dim(modelfit$data)[2]
+  pred <- predicted(modelfit)
+   res.df <- data.frame(ID = rep(seq_len(n_ts),
+                                 n_years),
+                        Time = sort(rep(seq_len(n_years), n_ts)), 
+                        mean = c(t(apply(pred, c(3, 4), mean))), 
+                        lo = c(t(apply(pred, c(3, 4), quantile, 0.025))),
+                        hi = c(t(apply(pred, c(3, 4), quantile, 0.975))), 
+                        y = c(modelfit$data))
+
+        res.df$ID <- new.names
+
+  res.df$year <-  rep(1972:2017, each=n_ts)
+
+  res.df$mean.res <- res.df$y - res.df$mean
+  res.df$lo.res <- res.df$y - res.df$lo
+  res.df$hi.res <- res.df$y - res.df$hi
+
+  index <- unique(res.df$ID)
+
+# using 25-yr rolling windows!
+
+    dfa.cor <- data.frame()
+
+    for(i in 1:length(index)){    
+      # i <- 1 
+      temp <- res.df %>%
+        filter(ID==index[i])
+
+       for(ii in 13:58){
+       #  ii <- 13
+          win <- temp[(ii-12):(ii+12),]
+          temp.cor <- data.frame(ID=unique(win$ID),
+                                 year=win$year[13], 
+                                 window=25,
+                                 mean.cor=cor(win$y, win$mean, use="p"),
+                                 n.cor=sum(!is.na(win$y)))
+          dfa.cor <- rbind(dfa.cor, temp.cor)
+       }  }
+
+    # remove correlations with < 10 obs
+dfa.cor <- dfa.cor %>%
+  filter(n.cor>=10)
+ dfa.cor$window <- as.factor(dfa.cor$window)
+
+dfa.cor <- na.omit(dfa.cor)
+
+# and plot 25-yr windows
+ggplot(filter(dfa.cor, window==25), aes(year, mean.cor)) + 
+  theme_bw() + theme(axis.title.x = element_blank()) +
+  geom_line(color=cb[3]) + facet_wrap(~ID, 
+        scales = "free_y", ncol=4) + xlab("Year") + ylab("Correlation")
+
+ggsave("goa biology dfa correlations 25 year windows all time series.png", height = 6.5, width=8, units="in")
 ```

### 2019-09-17 ewidata update - GOA.R

@@ -546,3 +546,183 @@ summary_table<-dfa_summary$summary
 capture.output(summary_table, file = paste0(name, "_summary.txt"))
 
 dev.off() 
+
+######################
+# one-trend models fit separately to first 15 and last 15 years of the TS
+sub_data <- goa.clim %>%
+  select(-ssh, -wind.stress, -salinity.20m) %>%
+  filter(year <= 1964) %>%
+  gather(key="code", value="value", -year)
+
+# ONE! trend model
+# with original data
+max_trends = 1
+name <- "GOA_clim_1_trend_original_data_1950_1964"
+
+# reshape data
+melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
+Y <- dcast(melted, code ~ year)
+names = Y$code
+Y = as.matrix(Y[,-which(names(Y) == "code")])
+
+# do the trend search
+set.seed(99)
+dfa_summary = find_dfa_trends(
+  y = Y,
+  kmax = min(max_trends, nrow(Y)),
+  iter = mcmc_iter,
+  compare_normal = FALSE,
+  variance = c("unequal", "equal"),
+  chains = mcmc_chains
+)
+saveRDS(dfa_summary, file = paste0(name, ".rds"))
+
+# Make default plots (currently work in progress)
+pdf(paste0(name, "_plots.pdf"))
+rotated = rotate_trends(dfa_summary$best_model)
+# trends
+print(plot_trends(rotated, years = as.numeric(colnames(Y))))
+# loadings
+print(plot_loadings(rotated, names = names))
+
+if(ncol(rotated$Z_rot_mean)==2) {
+  plot(rotated$Z_rot_mean[,1], rotated$Z_rot_mean[,2], col="white",
+       xlab="Loading 1", ylab = "Loading 2")
+  text(rotated$Z_rot_mean[,1], rotated$Z_rot_mean[,2], names, cex=0.3)
+  lines(c(-10,10),c(0,0))
+  lines(c(0,0), c(-10,10))
+}
+# predicted values with data
+print(plot_fitted(dfa_summary$best_model,names=names) + 
+        theme(strip.text.x = element_text(size = 6)))
+
+# table of AIC and std errors
+summary_table<-dfa_summary$summary
+capture.output(summary_table, file = paste0(name, "_summary.txt"))
+
+dev.off() 
+
+
+# last 15
+sub_data <- goa.clim %>%
[…]

### 2019-09-17 ewidata update - GOA.R
Finished loading comparison plot

@@ -348,6 +348,8 @@ options(mc.cores = parallel::detectCores())
 # find dfa trends with 2 or 3 candidate trends 
 # and original data (as in previous iteration of paper)
 # or expanded data (adding SSH and wind stress)
+goa.clim <- read.csv("updated goa climate data.csv")
+
 sub_data <- goa.clim %>%
   gather(key="code", value="value", -year)
 
@@ -700,8 +702,8 @@ loadings.2$era <- "2005-2019"
 
 
 # plot
-plot.load <- cbind(loadings.1, loadings.2) %>%
-  gather(-era)
+plot.load <- rbind(loadings.1, loadings.2) %>%
+  gather(key, value, -era)
 
 # rank <- tapply(plot.load$value, plot.load$key, mean)
 # plot.load$rank <- rank[match(plot.load$key, names(rank))]
@@ -718,11 +720,10 @@ plot.a <- ggplot(plot.load, aes(key, value)) +
   theme(axis.title.y = element_blank()) + ylab("Loading") +
   coord_flip() +
   ggtitle("a) Loadings")
-AL.p <- ggplot(slp, aes(slp.11, fill=era)) + 
-  geom_density(alpha=0.8) +
-  xlim(-700,700) + 
+
+ggplot(plot.load, aes(value, fill=era)) + 
+  geom_density(alpha=0.5) +
   scale_fill_manual(values=cb[c(6,8)]) +
   theme_bw() +
-  theme(legend.position = c(0.2,0.8), legend.title = element_blank()) +
-  xlab("Sea level pressure anomaly (Pa)") +
-  ggtitle("Aleutian Low SLPa distribution")
+  facet_wrap(~key, scales = "free_y") +
+  xlim(-1.5, 2.5) + geom_vline(xintercept = 0, lty=2)

### 2019-10-09 paper draft.Rmd
Trying to extract HMM results

@@
 set.seed(99)
-f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)
+f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 5)
 
@@
 plot_regime_model(f1$best_model)
+```
+
+```{r}
+# try to extract annual P of each state!
+    # gamma_tk <- rstan::extract(f1$best_model, pars = "gamma_tk")[[1]]
+    # mu_k <- rstan::extract(f1$best_model, pars = "mu_k")[[1]]
+    # l <- apply(gamma_tk, 2:3, quantile, probs = probs[[1]])
+    # u <- apply(gamma_tk, 2:3, quantile, probs = probs[[2]])
+    # med <- apply(gamma_tk, 2:3, quantile, probs = 0.5)
+
 ```

### 2019-11-19 updated plankton model.R
Updating plankton model

@@ -1,0 +1,67 @@
+library(tidyverse)
+library(ncdf4)
+library(chron)
+library(fields)
+library(maps)
+library(mapdata)
+library(zoo)
+library(gtools)
+
+# load updated ewidata!
+ewidata <- read.csv("ewidata.csv", row.names=1)
+
+# now load sewardline updates
+
+sew <- read.csv("sewardline.updates.csv")
+
+head(sew)
+cor(sew[,2:9])
+
+# and process
+# I think for now we should just use the biomass variable rather than biomass and abundance
+sew <- sew %>%
+  select(1:5) %>%
+  gather(key="code", value="value", -year)
+
+ggplot(sew, aes(log(value))) +
+  theme_bw() +
+  geom_histogram(bins=12) +
+  facet_wrap(~code, scales="free")
+
+# log transform
+sew$value <- log(sew$value)
+
+# pull old Sewardline data and replace
+drop <- grep("SEWARD", ewidata$code)
+ewidata <- ewidata[-drop,]
+
+# and replace with new
+sew$system <- sew$subtype <- NA
+
+# and add "SEWARD_" to codes
+sew$code <- paste("SEWARD_", sew$code, sep="")
+ewidata <- rbind(ewidata, sew)
+
+# now add primary productivity data
+pp <- read.csv("GOA-primary-production.csv")
+
+pp <- pp %>%
+  gather(key="code", value="value", -year)
+
+ggplot(pp, aes(value)) +
+  theme_bw() +
+  geom_histogram(bins=8) +
+  facet_wrap(~code, scales="free")
+
+# I think we can live with that distribution 
+# - although I should ask Eric/Sean if the bimodal distribution in duration is a problem for the DFA models
+
+pp$code <- paste("PP_", pp$code, sep="")
+
+pp$system <- pp$subtype <- NA
+
+ewidata <- rbind(ewidata, pp)
+
+# now updated CPR data
+
+filter(ewidata, code %in% c("copepod", "mesozoo", "diatoms"))

### 2019-11-19 paper draft.Rmd
Changed # of HMM states to 5

--- 2019-09-17 paper draft.Rmd
@@
 
-Test for regime probability. Use n=3 maximum states.
+Test for regime probability. Use n=5 maximum states.
 
@@
 plot_regime_model(f1$best_model)
+```
+
+```{r}
+# try to extract annual P of each state!
+
+kk <- plot_regime_model(f1$best_model, plot_prob_indices=c(2))
+
+clim.regimes <- data.frame(year=1950:2019, probability=kk$data$median)
+
+ggplot(clim.regimes, aes(year, probability)) +
+  theme_bw() +
+  geom_line()
+
+# may be promising to plot all of the regime results together!
+
+clim.regimes$data.set <- "Climate 1950-2019"
 ```
@@
 plot_regime_model(f1$best_model)
+
+# and combined plot
+
+kk <- plot_regime_model(f1$best_model, plot_prob_indices=c(2))
+
+biol.regimes <- data.frame(year=1972:2017, probability=kk$data$median)
+
+biol.regimes$data.set <- "Biology 1972-2017"
+
+plot.regimes <- rbind(clim.regimes, biol.regimes)
+
+ggplot(plot.regimes, aes(year, probability, color=data.set)) +
+  theme_bw() +
+  geom_line()
 
@@
 ggsave("goa biology era sst pdo slopes.png", width=6, height=4, units="in")
+
+# make a SST-only plot
+ggplot(melt.mod, aes(x=value, fill=era)) +
+  theme_bw() + 
+  geom_density(alpha=0.6) +
+  geom_vline(xintercept = 0, lty=2) +
+  scale_fill_manual(values = cb[c(4, 6, 2)]) +
+  theme(legend.title = element_blank(), legend.position = c(0.8, 0.8)) +
+  xlab("Slope") +
+  xlim(-5,12)
+
+ggsave("goa biology era sst slopes.png", width=5, height=4, units="in")
+
 ```
@@
 plot_regime_model(f1$best_model)
+
+# and combined plot
+
+kk <- plot_regime_model(f1$best_model, plot_prob_indices=c(1))
+
+ichthyo.regimes <- data.frame(year=plot.trend$year, probability=kk$data$median)
+
+ichthyo.regimes$data.set <- "Ichthyoplankton 1980-2015"
+
+plot.regimes <- rbind(clim.regimes, biol.regimes, ichthyo.regimes)
+
+ggplot(plot.regimes, aes(year, probability, color=data.set)) +
+  theme_bw() +
+  geom_line()
 
@@
 
+# and combined plot
+
+kk <- plot_regime_model(f1$best_model, plot_prob_indices=c(1))
+
+zoop.regimes <- data.frame(year=plot.trend$year, probability=kk$data$median)
+
+zoop.regimes$data.set <- "Zooplankton 1997-2016"
+
+plot.regimes <- rbind(clim.regimes, biol.regimes, ichthyo.regimes, zoop.regimes)
+
+plot.regimes$plot.order <- ifelse(plot.regimes$data.set=="Climate 1950-2019", 1, 
+                                  ifelse(plot.regimes$data.set=="Biology 1972-2017", 2,
+                                  ifelse(plot.regimes$data.set=="Ichthyoplankton 1980-2015", 3,
+                                  4)))
+plot.regimes$data.set <- reorder(plot.regimes$data.set, plot.regimes$plot.order)
+
+ggplot(plot.regimes, aes(year, probability, color=data.set)) +
+  theme_bw() +
+  geom_line() +
+  theme(axis.title.x = element_blank(), legend.title = element_blank(), legend.position = 'top') + 
+  scale_color_manual(values = cb[c(6,4,2,7,8)], guide=guide_legend(nrow=2))
+
+ggsave("regime probabilities combined.png", width = 5, height = 3.5, units = "in")
+
 ```

### 2019-11-19 ewidata update - GOA biology.R
Updated DFA and era-dependent loadings

@@ -139,6 +139,10 @@ GOA <- rbind(GOA, xtra1, xtra2)
 # and save
 write.csv(GOA, "updated GOA biology data.csv", row.names = F)
 
+# reload
+GOA <- read.csv("updated GOA biology data.csv")
+
+
 # and check one more time
 ggplot(GOA, aes(year, value)) +
   theme_bw() +
@@ -146,6 +150,7 @@ ggplot(GOA, aes(year, value)) +
   geom_point() +
   facet_wrap(~code, scales="free_y")
 
+##########################################
 # now fit models!
 library(Rcpp)
 library(ewidata)
@@ -158,6 +163,10 @@ library(gtools)
 # devtools::install_github("fate-ewi/bayesdfa", force=T)
 
 library(bayesdfa)
+
+# reload data if needed
+
+GOA <- read.csv("updated GOA biology data.csv")
 
 mcmc_iter = 4000
 max_trends = 1 # changing to 1!
@@ -214,4 +223,180 @@ capture.output(summary_table, file = paste0(name, "_summary.txt"))
 
 dev.off()
 
+############################################
+# fit to the first and last 15 years, as for climate data
+######################
+# one-trend models fit separately to first 15 and last 15 years of the TS
+sub_data <- GOA %>%
+  select(year, code, value) %>%
+  filter(year <= 1986) 
 
+# ONE! trend model
+# with original data
+max_trends = 1
+name <- "GOA_biol_1_trend_1972_1986"
+
+# reshape data
+melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
+Y <- dcast(melted, code ~ year)
+names = Y$code
+Y = as.matrix(Y[,-which(names(Y) == "code")])
+
+# do the trend search
+set.seed(99)
+dfa_summary = find_dfa_trends(
+  y = Y,
+  kmax = min(max_trends, nrow(Y)),
+  iter = mcmc_iter,
+  compare_normal = FALSE,
+  variance = c("unequal", "equal"),
+  chains = mcmc_chains
+)
+saveRDS(dfa_summary, file = paste0(name, ".rds"))
+
+# Make default plots (currently work in progress)
+pdf(paste0(name, "_plots.pdf"))
+rotated = rotate_trends(dfa_summary$best_model)
+# trends
[…]

### 2019-11-19 updated plankton model.R
Processing data & re-running DFA

@@ -1,11 +1,14 @@
-library(tidyverse)
-library(ncdf4)
-library(chron)
-library(fields)
-library(maps)
-library(mapdata)
-library(zoo)
+library(Rcpp)
+library(ewidata)
+library(knitr)
+library(reshape2)
+library(rstan)
+library(tibble)
+library(dplyr)
 library(gtools)
+# devtools::install_github("fate-ewi/bayesdfa", force=T)
+
+library(bayesdfa)
 
 # load updated ewidata!
 ewidata <- read.csv("ewidata.csv", row.names=1)
@@ -28,15 +31,13 @@ ggplot(sew, aes(log(value))) +
   geom_histogram(bins=12) +
   facet_wrap(~code, scales="free")
 
-# log transform
-sew$value <- log(sew$value)
-
 # pull old Sewardline data and replace
 drop <- grep("SEWARD", ewidata$code)
 ewidata <- ewidata[-drop,]
 
 # and replace with new
-sew$system <- sew$subtype <- NA
+sew$system <- "WGOA"
+sew$subtype <- NA
 
 # and add "SEWARD_" to codes
 sew$code <- paste("SEWARD_", sew$code, sep="")
@@ -58,10 +59,133 @@ ggplot(pp, aes(value)) +
 
 pp$code <- paste("PP_", pp$code, sep="")
 
-pp$system <- pp$subtype <- NA
+pp$system <- "EGOA"
+pp$subtype <- NA
 
 ewidata <- rbind(ewidata, pp)
+# 
+# # now updated CPR data
+# 
+# filter(ewidata, code %in% c("copepod", "mesozoo", "diatoms"))
 
-# now updated CPR data
+# GoA biology
 
-filter(ewidata, code %in% c("copepod", "mesozoo", "diatoms"))
+codes <- unique(ewidata$code)
+
+# limit to GOA
+GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
+unique(GOA$code)
+
+# drop salmon
+drop <- grep("PI", GOA$code)
+GOA <- GOA[-drop,]
+
+drop <- grep("CO", GOA$code)
+GOA <- GOA[-drop,]
+
+# drop climate
+drop <- grep("AKCLIM", GOA$code)
+GOA <- GOA[-drop,]
+
+unique(GOA$code)
+
+# drop Pavlof and Chiniak and herring and kittiwakes
[…]

### 2019-11-20 ewidata update - GOA.R
Updated and ran three separate plankton models

@@ -713,14 +713,6 @@ plot.load <- rbind(loadings.1, loadings.2) %>%
 # set pallette
 cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 
-plot.a <- ggplot(plot.load, aes(key, value)) +
-  theme_bw() +
-  geom_violin(fill=cb[3]) +
-  geom_hline(yintercept = 0) +
-  theme(axis.title.y = element_blank()) + ylab("Loading") +
-  coord_flip() +
-  ggtitle("a) Loadings")
-
 ggplot(plot.load, aes(value, fill=era)) + 
   geom_density(alpha=0.5) +
   scale_fill_manual(values=cb[c(6,8)]) +

### 2019-11-20 updated plankton model.R

@@ -18,15 +18,24 @@ ewidata <- read.csv("ewidata.csv", row.names=1)
 sew <- read.csv("sewardline.updates.csv")
 
 head(sew)
-cor(sew[,2:9])
+
+# calculate size time series as per Russ' suggestion
+sew$calanoid.size.spring <- sew$LB_Calanoid.Spr / sew$Calanoid.A.spr
+sew$euphaus.size.spring <- sew$LB_Euphaus.Spr / sew$Euphausiid.A.spr
+sew$calanoid.size.sept <- sew$LB_Calanoid.Sep / sew$Calanoid.A.Sept
+sew$euphaus.size.sept <- sew$LB_Euphaus.Sep / sew$Euphausiid.A.Sep
+
+
+cor(sew[,2:13])
 
 # and process
 # I think for now we should just use the biomass variable rather than biomass and abundance
+# and also adding size!
 sew <- sew %>%
-  select(1:5) %>%
+  select(1:5, 10:13) %>%
   gather(key="code", value="value", -year)
 
-ggplot(sew, aes(log(value))) +
+ggplot(sew, aes(value)) +
   theme_bw() +
   geom_histogram(bins=12) +
   facet_wrap(~code, scales="free")
@@ -65,8 +74,57 @@ pp$subtype <- NA
 ewidata <- rbind(ewidata, pp)
 # 
 # # now updated CPR data
-# 
+
+cpr <- read.csv("CPR.2019.csv")
+names(cpr)[1] <- "year"
+cor(cpr[,2:6])
+
+cpr <- cpr %>%
+  gather(key="code", value, -year)
+
+ggplot(cpr, aes(value)) +
+  theme_bw() +
+  geom_histogram(bins=8) +
+  facet_wrap(~code, scales="free")
+# so these are anomalies from log-transformed means and do not need to be transformed again
+
+# append "CPR" to the names
+cpr$code <- paste("CPR_", cpr$code, sep="")
+
+cpr$system <-  "EGOA"
+cpr$subtype <- NA
+
+# drop existing CPR
+
+ewidata <- ewidata %>%
+  filter(code != "copepod") %>%
+  filter(code != "mesozoo") %>%
+  filter(code != "diatoms")
+
 # filter(ewidata, code %in% c("copepod", "mesozoo", "diatoms"))
+
+ewidata <- rbind(ewidata, cpr)
+
+# and finally, ICY strait updates
+# first, drop existing ICY data
+drop <- grep("ICY", ewidata$code)
[…]

### 2019-11-20 updated plankton model.R
Fixed bug dropping icy copepods by mistake

@@ -126,74 +126,54 @@ cor(check[,4:9])
 # combine with ewidata
 ewidata <- rbind(ewidata, icy)
 
+# and save
+write.csv(ewidata, "/Users/MikeLitzow/Documents/R/FATE/ewidata/inst/extdata/ewidata_11.20.19", row.names = F)
+write.csv(ewidata, "ewidata.plankton.updated.csv", row.names=F)
+
+
 # GoA biology
 
 codes <- unique(ewidata$code)
-
-# limit to GOA
-GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
-unique(GOA$code)
-
-# drop salmon
-drop <- grep("PI", GOA$code)
-GOA <- GOA[-drop,]
-
-drop <- grep("CO", GOA$code)
-GOA <- GOA[-drop,]
-
-# drop climate
-drop <- grep("AKCLIM", GOA$code)
-GOA <- GOA[-drop,]
-
-unique(GOA$code)
-
-# drop Pavlof and Chiniak and herring and kittiwakes
-drop <- grep("PAV", GOA$code)
-GOA <- GOA[-drop,]
-
-drop <- grep("CHI", GOA$code)
-GOA <- GOA[-drop,]
-
-drop <- grep("HER", GOA$code)
-GOA <- GOA[-drop,]
-
-drop <- grep("BLKI", GOA$code)
-GOA <- GOA[-drop,]
+GOA <- ewidata
 
 # drop total icy zoops!
 drop <- grep("ICY.ZOOP.TOTDEN", GOA$code)
 GOA <- GOA[-drop,]
 
-# and drop ichthyoplankton
-drop <- grep("ICH", GOA$code)
-GOA <- GOA[-drop,]
-
-# # and try dropping PP duration, as this is bimodal and might give trouble with the model 
-# # returning bimodal loadings
-# drop <- grep("_dur", GOA$code)
-# GOA <- GOA[-drop,]
-
-unique(GOA$code)
-
-# now check for outliers!
-ggplot(GOA, aes(x=year, y = value))  + 
-  theme_bw() + 
-  geom_line() +
-  facet_wrap(~code, scales = "free_y")  + xlab("Year") +
-  ylab("Value")
-
-ggplot(GOA, aes(value))  + 
-  theme_bw() + 
-  geom_histogram(fill="grey", color="black") +
-  facet_wrap(~code, scales = "free")  
-
-# log transform ICY and Seward
[…]

### 2019-11-22 paper draft.Rmd
Tweaks to plankton processing

@@
 Now the zoops model.
-
-```{r, echo=F}
-GOA.zoop <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_zoop_7.19.18.one.trend.rds") 
-GOA.zoop$summary
+These are now fit as three models- one for Seward Line, one for Icy Strait, one for large-scale (CPR & PP)
+```{r, echo=F}
+seward <- readRDS("Sewardline_11.19.19.one.trend.rds") 
+seward$summary
 ```
@@
 # load updated ewidata!
-ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
-
-# limit to GOA
-GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
-unique(GOA$code)
-
-# drop salmon
-drop <- grep("PI", GOA$code)
-GOA <- GOA[-drop,]
-
-drop <- grep("CO", GOA$code)
-GOA <- GOA[-drop,]
-
-# drop climate
-drop <- grep("AKCLIM", GOA$code)
-GOA <- GOA[-drop,]
-
-unique(GOA$code)
-
-# drop Pavlof and Chiniak and herring and kittiwakes
-drop <- grep("PAV", GOA$code)
-GOA <- GOA[-drop,]
-
-drop <- grep("CHI", GOA$code)
-GOA <- GOA[-drop,]
-
-drop <- grep("HER", GOA$code)
-GOA <- GOA[-drop,]
-
-drop <- grep("BLKI", GOA$code)
-GOA <- GOA[-drop,]
-
-# drop total icy zoops!
-drop <- grep("ICY.ZOOP.TOTDEN", GOA$code)
-GOA <- GOA[-drop,]
-
-# and drop ichthyoplankton
-drop <- grep("ICH", GOA$code)
-GOA <- GOA[-drop,]
-
-unique(GOA$code)
+ewidata <- as.tibble(read.csv("ewidata.plankton.updated.csv"))
+
+keep <- grep("SEWARD", ewidata$code)
+seward.dat <- ewidata[keep,]
+unique(seward.dat$code)
+
 # recover the years and names for plots!
-sub_data = GOA
+sub_data = seward.dat
 
@@
 
-rotated = rotate_trends(GOA.zoop$best_model)
-
-new.names <- c("CPR.copepods", "CPR.diatoms", "Icy.Str.amphipods", "Icy.Str.euphausids", "Icy.Str.gastropods", "CPR.mesozooplankton", "Seward.A.calanoid", "Seward.A.euphausid", "Seward.B.calanoid", "Seward.B.euphausid", "Seward.LA.euphausid", "Seward.LB.calanoid", "Seward.LB.euphausid", "Seward.calanoid.size", "Seward.euphausid.size", "Seward.LA.calanoid")
+rotated = rotate_trends(seward$best_model)
+
+new.names <- c("Calanoid.size.Sept", "Calanoid.size.May", "Euphausid.size.Sept", "Euphausid.size.May",
+               "Calanoid.biomass.Sept", "Calanoid.biomass.May", "Euphausid.biomass.Sept", "Euphausid.biomass.May")
 
@@
   coord_flip() +
-  ggtitle("d) Zooplankton loadings") +
+  ggtitle("d) Sewardline loadings") +
   theme(title = element_text(size=8))
@@
   geom_hline(yintercept = 0) + ylab("Trend value") + xlab("Year") +
-  ggtitle("e) Zooplankton trend") +
+  ggtitle("e) Sewardline trend") +
   theme(title = element_text(size=8))
@@
 
-zoop.regimes <- data.frame(year=plot.trend$year, probability=kk$data$median)
-
-zoop.regimes$data.set <- "Zooplankton 1997-2016"
+seward.regimes <- data.frame(year=plot.trend$year, probability=kk$data$median)
+
+seward.regimes$data.set <- "Sewardline 1998-2018"
 
@@
 
-years <- as.numeric(colnames(GOA.zoop$best_model$data))
+years <- as.numeric(colnames(seward$best_model$data))
 
@@
   ylab("Density") +
-  ggtitle(("f) Zooplankton slope on SST")) +
+  ggtitle(("f) Sewardline slope on SST")) +
   theme(title = element_text(size=8))
@@
 
+Now the icy strait model.
+
+```{r, echo=F}
+icy <- readRDS("ICY_11.20.19.one.trend.rds") 
+icy$summary
+```
+
+```{r, echo=F}
+
+# load updated ewidata!
+ewidata <- as.tibble(read.csv("ewidata.plankton.updated.csv"))
+
+keep <- grep("ICY", ewidata$code)
+icy.dat <- ewidata[keep,]
+unique(icy.dat$code)
+
+# drop total density
+drop <- grep("TOTDEN", icy.dat$code)
+icy.dat <- icy.dat[-drop,]
+
+# recover the years and names for plots!
+sub_data = icy.dat
+
+melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
+Y <- dcast(melted, code ~ year)
+names = Y$code
+Y = as.matrix(Y[,-which(names(Y) == "code")])
+
+rotated = rotate_trends(icy$best_model)
+
+new.names <- c("Amphiphod", "Large.copepod", "Small.copepod", "Euphausid",
+               "Gastropod")
+
+# back to the bespoke code!
+loadings.1 <- as.data.frame(rotated$Z_rot[,,1])
+names(loadings.1)  <- new.names
+
+# drop columns that don't meet threshold
+# first, figure out the proportion above/below 0
+f <- function(x) max(sum(x<0)/length(x), sum(x>0)/length(x))
+
+prop1 <- apply(loadings.1, 2, f)
+prop1
+keep <- prop1 > 0.95
+
+loadings.1 <- loadings.1[,keep]
+
+# plot
+# AND THERE IS ONLY ONE TS WITH > 0.95 LOADINGS!
+plot.load <- loadings.1 %>%
+  gather()
+
+rank <- tapply(plot.load$value, plot.load$key, mean)
+plot.load$rank <- rank[match(plot.load$key, names(rank))]
+
+plot.load$key <- reorder(plot.load$key, plot.load$rank)
+
+# set pallette
+cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+plot.d <- ggplot(plot.load, aes(key, value)) +
+  theme_bw() +
+  geom_violin(fill=cb[3]) +
+  geom_hline(yintercept = 0) +
+  theme(axis.title.y = element_blank()) + ylab("Loading") +
+  coord_flip() +
+  ggtitle("d) Sewardline loadings") +
+  theme(title = element_text(size=8))
+
+plot.d
+```
+
+Now the large-scale!
+
+```{r, echo=F}
+large <- readRDS("PP_CPR_11.20.19.one.trend.rds") 
+large$summary
+```
+
+```{r, echo=F}
+
+# load updated ewidata!
+ewidata <- as.tibble(read.csv("ewidata.plankton.updated.csv"))
+
+keep <- c(grep("PP", ewidata$code), grep("CPR", ewidata$code))
+large.dat <- ewidata[keep,]
+unique(large.dat$code)
+
+# recover the years and names for plots!
+sub_data = large.dat
+
+melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
+Y <- dcast(melted, code ~ year)
+names = Y$code
+Y = as.matrix(Y[,-which(names(Y) == "code")])
+
+rotated = rotate_trends(large$best_model)
+
+new.names <- c("Copepod.size", "Diatom.abundance", "Large.copepod.abundance", "Mesozoop.abundance",
+               "Small.copepod.abundance", "Bloom.amplitude", "Bloom.duration", "Bloom.start")
+
+# back to the bespoke code!
+loadings.1 <- as.data.frame(rotated$Z_rot[,,1])
+names(loadings.1)  <- new.names
+
+# drop columns that don't meet threshold
+# first, figure out the proportion above/below 0
+f <- function(x) max(sum(x<0)/length(x), sum(x>0)/length(x))
+
+prop1 <- apply(loadings.1, 2, f)
+prop1
+keep <- prop1 > 0.95
+
+loadings.1 <- loadings.1[,keep]
+
+# plot
+plot.load <- loadings.1 %>%
+  gather()
+
+rank <- tapply(plot.load$value, plot.load$key, mean)
+plot.load$rank <- rank[match(plot.load$key, names(rank))]
+
+plot.load$key <- reorder(plot.load$key, plot.load$rank)
+
+# set pallette
+cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+plot.g <- ggplot(plot.load, aes(key, value)) +
+  theme_bw() +
+  geom_violin(fill=cb[3]) +
+  geom_hline(yintercept = 0) +
+  theme(axis.title.y = element_blank()) + ylab("Loading") +
+  coord_flip() +
+  ggtitle("g) Large-scale loadings") +
+  theme(title = element_text(size=8))
+
+plot.g
+```
+
+```{r, echo=F}
+# trends
+plot.trend <- data.frame(year=as.numeric(as.character(colnames(Y))), trend=as.vector(rotated$trends_mean),
+                         lo=as.vector(rotated$trends_lower),
+                         hi=as.vector(rotated$trends_upper))
+
+plot.h <- ggplot(plot.trend, aes(year, trend)) +
+  theme_bw() +
+  geom_line(color=cb[3]) +
+  geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
+  geom_hline(yintercept = 0) + ylab("Trend value") + xlab("Year") +
+  ggtitle("h) Large-scale trend") +
+  theme(title = element_text(size=8))
+
+plot.h
+```
+
+Now test for SST effect.
+```{r, include=F}
+#  load sst as the explanatory variable.
+ewidata <- read.csv("updated goa climate data.csv", row.names=1)
+
+keep <- grep("win.sst", colnames(ewidata))
+
+clim.var <- ewidata[,keep] %>%
+  mutate(year=1950:2019) %>%
+  gather(key, value, -year)
+
+win.sst <- tapply(clim.var$value, clim.var$year, mean)
+
+years <- as.numeric(colnames(seward$best_model$data))
+
+# set up climate data
+cli_dat <- data.frame(Year=as.numeric(names(win.sst)),SST=win.sst)
+
+cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))
+
+# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
+
+cli_dat$slope <- 1
+#cli_dat$slope[cli_dat$Year %in% 1997:2013] <- 2
+#cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
+cli_dat$slope <- as.factor(cli_dat$slope)
+
+# rename rotated
+r <- rotated
+
+Y = rnorm(length(cli_dat$slope))
+cli_dat = as.data.frame(model.matrix(Y ~ Year + SST, data=cli_dat))
+# replace ':' in name for formula
+names(cli_dat) = gsub(":","",names(cli_dat))
+# remove named intercept
+cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
+
+set.seed(99)
+mod = trend_lm(rotated_modelfit = r, all_years = years,
+  selected_years = years, climate_dat=cli_dat)
+
+summary(mod)
+
+# get 95% CIs
+x <- sort(mod[,2])
+print(names(mod)[2])
+print(paste("median: ", median(x)))
+print(paste("LCI: ", x[(0.025*25000)]))
+print(paste("UCI: ", x[(0.975*25000)]))
+
+plot.i <- ggplot(mod, aes(x=SST)) +
+  theme_bw() +
+  geom_density(fill=cb[3]) + 
+  geom_hline(yintercept=0) + 
+  geom_vline(xintercept = 0, lty=2) + 
+  xlab("Slope") + 
+  ylab("Density") +
+  ggtitle(("i) Large-scale slope on SST")) +
+  theme(title = element_text(size=8))
+
+```
+
+Look at era-specific effects.
+```{r, include=F}
+# first, define the function
+# increasing to 4000 iterations!
+trend_lm <- function(rotated_modelfit,
+  all_years, selected_years,
+  climate_dat = NULL,
+  trend_number = 1, samples = 50,
+  ITER = 4000, CHAINS = 1, n_eff = 50, rhat = 1.1, show_plot = TRUE) { # increasing from 1000 to 4000 iterations to help plankton model converge
+
+  total_mcmc <- dim(rotated_modelfit$trends)[1]
+  draws <- sample(seq_len(total_mcmc), samples)
+
+  yrs_i <- which(all_years %in% selected_years)
+  climate_dat <- dplyr::filter(climate_dat, Year %in% selected_years)
+
+  predictor_name = names(climate_dat)[which(names(climate_dat) != "Year")]
+  formula_str = paste("dfa_trend_draw ~", predictor_name[1])
+  if(length(predictor_name) > 1) {
+    for(i in 2:length(predictor_name)) {formula_str = paste(formula_str, predictor_name[i], sep=" + ")}
+  }
+  formula <- as.formula(formula_str)
+
+  post <- list()
+  for (i in seq_along(draws)) {
+    dd = climate_dat
+    dd$dfa_trend_draw = rotated_modelfit$trends[draws[i], trend_number, yrs_i]
+    m_temp <- stan_glm(formula, data = dd,
+      iter = ITER, chains = CHAINS)
+    assert_that(all(m_temp$stan_summary[,"Rhat"] < rhat))
+    assert_that(all(m_temp$stan_summary[,"n_eff"] > n_eff))
+    post[[i]] <- as.data.frame(as.matrix(m_temp))
+  }
+  post_all <- dplyr::bind_rows(post)
+
+  if (show_plot) {
+    #par(mfrow = c(1, 2))
+    #plot(post_all[,predictor], type = "l", col = "#00000080", lwd = 0.5)
+    #abline(v = seq(0, ITER * length(draws), ITER), col = "#FF000050")
+    #plot(density(post_all[,predictor]))
+    #abline(v = 0, col = "#00000050")
+  }
+  invisible(post_all)
+}
+
+#  load sst as the explanatory variable.
+
+ewidata <- read.csv("updated goa climate data.csv", row.names=1)
+
+keep <- grep("win.sst", colnames(ewidata))
+
+clim.var <- ewidata[,keep] %>%
+  mutate(year=1950:2019) %>%
+  gather(key, value, -year)
+
+win.sst <- tapply(clim.var$value, clim.var$year, mean)
+
+years <- as.numeric(colnames(large$best_model$data)) # these are the years for the biology model - 1972:2016
+
+# set up climate data
+cli_dat <- data.frame(Year=as.numeric(names(win.sst)),SST=win.sst)
+
+cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))
+
+# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
+
+cli_dat$slope <- 1
+cli_dat$slope[cli_dat$Year %in% 2014:2019] <- 2
+cli_dat$slope <- as.factor(cli_dat$slope)
+
+# rename rotated
+r <- rotated
+
+Y = rnorm(length(cli_dat$slope))
+cli_dat = as.data.frame(model.matrix(Y ~ Year + SST:slope, data=cli_dat))
+# replace ':' in name for formula
+names(cli_dat) = gsub(":","",names(cli_dat))
+# remove named intercept
+cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
+
+set.seed(99)
+mod = trend_lm(rotated_modelfit = r, all_years = 1998:2019,
+  selected_years = 1998:2019, climate_dat=cli_dat)
+
+names(mod)[2:3] <- c("1998-2013", "2014-2019")
+summary(mod)
+
+# get 95% CIs
+
+for(i in 2:3){
+x <- sort(mod[,i])
+print(names(mod)[i])
+print(paste("median: ", median(x)))
+print(paste("LCI: ", x[(0.025*25000)]))
+print(paste("UCI: ", x[(0.975*25000)]))
+}
+
+melt.mod <- melt(mod, measure.vars=c("1998-2013", "2014-2019"))
+names(melt.mod)[3] <- "era"
+melt.mod$covar <- "SST"
+
+# make a SST-only plot
+ggplot(melt.mod, aes(x=value, fill=era)) +
+  theme_bw() + 
+  geom_density(alpha=0.6) +
+  geom_vline(xintercept = 0, lty=2) +
+  scale_fill_manual(values = cb[c(6, 2)]) +
+  theme(legend.title = element_blank(), legend.position = c(0.8, 0.8)) +
+  xlab("Slope") #+
+  #xlim(-5,12)
+
+ggsave("large area era sst slopes.png", width=5, height=4, units="in")
+
+```
+
 And combine into a single plot.
@@
 ```{r}
-png("combined short time series plots.png", 8,5, units="in", res=300)
-
-ggpubr::ggarrange(plot.a, plot.b, plot.c, plot.d, plot.e, plot.f,
-          ncol=3, widths=c(1, 1, 0.8), nrow=2)
+png("combined updated plankton time series plots.png", 8,7.5, units="in", res=300)
+
+ggpubr::ggarrange(plot.a, plot.b, plot.c, plot.d, plot.e, plot.f, plot.g, plot.h, plot.i,
+          ncol=3, widths=c(1, 1, 0.8), nrow=3)
 
@@
 ```
+
+Or, much better! I ran a single model for the entire plankton TS and this appears to work...
+
+```{r, echo=F}
+all <- readRDS("all_plankton_11.20.19.one.trend.rds") 
+all$summary
+```
+
+```{r, echo=F}
+
+# load updated ewidata!
+ewidata <- as.tibble(read.csv("ewidata.plankton.updated.csv"))
+
+GOA <- ewidata
+
+# drop total icy zoops!
+drop <- grep("ICY.ZOOP.TOTDEN", GOA$code)
+GOA <- GOA[-drop,]
+
+# select ICY, PP, Seward, CPR, and ichthyo!
+keep <- c(grep("ICY", GOA$code), grep("PP_", GOA$code), grep("CPR", GOA$code), grep("SEWAR", GOA$code), grep("ICH", GOA$code))
+GOA <- GOA[keep,]
+
+# recover the years and names for plots!
+sub_data = GOA
+
+melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
+Y <- dcast(melted, code ~ year)
+names = Y$code
+Y = as.matrix(Y[,-which(names(Y) == "code")])
+
+rotated = rotate_trends(all$best_model)
+
+new.names <- c("CPR copepod size", "CPR diatom", "CPR large copepod", "CPR mesozooplankton", "CPR small copepod", "Ichthyo. - sand lance", 
+               "Ichthyo. - ronquils", "Ichthyo. - pollock", "Ichthyo. - Pacific cod", "Ichthyo. - flathead sole", "Ichthyo. - rock sole", "Ichthyo. - rockfishes", "Icy Strait amphipod",
+               "Icy Strait large copepod", "Icy Strait small copepod", "Icy Strait euphausid", "Icy Strait gastropod", "Bloom amplitude", "Bloom duration", "Bloom start", "Seward calanoid size Sept.",
+               "Seward calanoid size May", "Seward euphausid size Sept.", "Seward euphausid size May", "Seward calanoid Sept.", "Seward calanoid May", "Seward euphausid Sept.", "Seward euphausid May")
+
+# and check
+check <- data.frame(names, new.names)
+check
+
+# back to the bespoke code!
+loadings.1 <- as.data.frame(rotated$Z_rot[,,1])
+names(loadings.1)  <- new.names
+
+# drop columns that don't meet threshold
+# first, figure out the proportion above/below 0
+f <- function(x) max(sum(x<0)/length(x), sum(x>0)/length(x))
+
+prop1 <- apply(loadings.1, 2, f)
+prop1
+keep <- prop1 > 0.95
+
+loadings.1 <- loadings.1[,keep]
+
+# plot
+
+plot.load <- loadings.1 %>%
+  gather()
+
+rank <- tapply(plot.load$value, plot.load$key, mean)
+plot.load$rank <- rank[match(plot.load$key, names(rank))]
+
+plot.load$key <- reorder(plot.load$key, plot.load$rank)
+
+# set pallette
+cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+plot.a <- ggplot(plot.load, aes(key, value)) +
+  theme_bw() +
+  geom_violin(fill=cb[3], color=NA) +
+  geom_hline(yintercept = 0) +
+  theme(axis.title.y = element_blank()) + ylab("Loading") +
+  coord_flip() +
+  ggtitle("a) Plankton loadings") +
+  theme(title = element_text(size=8)) +
+  ylim(-2.5,2.5)
+
+plot.a
+```
+
+And trend.
+
+```{r, echo=F}
+# trends
+plot.trend <- data.frame(year=as.numeric(as.character(colnames(Y))), trend=as.vector(rotated$trends_mean),
+                         lo=as.vector(rotated$trends_lower),
+                         hi=as.vector(rotated$trends_upper))
+
+plot.b <- ggplot(plot.trend, aes(year, trend)) +
+  theme_bw() +
+  geom_line(color=cb[3]) +
+  geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
+  geom_hline(yintercept = 0) + ylab("Trend value") + xlab("Year") +
+  ggtitle("b) Plankton trend") +
+  theme(title = element_text(size=8))
+
+plot.b
+```
+
+Now the slope on SST.
+
+```{r, include=F}
+#  load sst as the explanatory variable.
+ewidata <- read.csv("updated goa climate data.csv", row.names=1)
+
+keep <- grep("win.sst", colnames(ewidata))
+
+clim.var <- ewidata[,keep] %>%
+  mutate(year=1950:2019) %>%
+  gather(key, value, -year)
+
+win.sst <- tapply(clim.var$value, clim.var$year, mean)
+
+years <- as.numeric(colnames(all$best_model$data))
+
+# set up climate data
+cli_dat <- data.frame(Year=as.numeric(names(win.sst)),SST=win.sst)
+
+cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))
+
+# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
+
+cli_dat$slope <- 1
+#cli_dat$slope[cli_dat$Year %in% 1997:2013] <- 2
+#cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
+cli_dat$slope <- as.factor(cli_dat$slope)
+
+# rename rotated
+r <- rotated
+
+Y = rnorm(length(cli_dat$slope))
+cli_dat = as.data.frame(model.matrix(Y ~ Year + SST, data=cli_dat))
+# replace ':' in name for formula
+names(cli_dat) = gsub(":","",names(cli_dat))
+# remove named intercept
+cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
+
+set.seed(99)
+mod = trend_lm(rotated_modelfit = r, all_years = years,
+  selected_years = years, climate_dat=cli_dat)
+
+summary(mod)
+
+# get 95% CIs
+x <- sort(mod[,2])
+print(names(mod)[2])
+print(paste("median: ", median(x)))
+print(paste("LCI: ", x[(0.025*25000)]))
+print(paste("UCI: ", x[(0.975*25000)]))
+
+plot.c <- ggplot(mod, aes(x=SST)) +
+  theme_bw() +
+  geom_density(fill=cb[3]) + 
+  geom_hline(yintercept=0) + 
+  geom_vline(xintercept = 0, lty=2) + 
+  xlab("Slope") + 
+  ylab("Density") +
+  ggtitle(("c) Plankton slope on SST")) +
+  theme(title = element_text(size=8))
+
+plot.c
+```
+
+And combine into a single plot.
+
+```{r}
+png("combined all plankton time series plots.png", 8, 2.5, units="in", res=300)
+
+ggpubr::ggarrange(plot.a, plot.b, plot.c,
+          ncol=3, widths=c(1, 1, 0.8), nrow=1)
+
+dev.off()
+```
+
+Bimodal - evidence of nonstationary slope?
+
+```{r}
+
+#  load sst as the explanatory variable.
+
+ewidata <- read.csv("updated goa climate data.csv", row.names=1)
+
+keep <- grep("win.sst", colnames(ewidata))
+
+clim.var <- ewidata[,keep] %>%
+  mutate(year=1950:2019) %>%
+  gather(key, value, -year)
+
+win.sst <- tapply(clim.var$value, clim.var$year, mean)
+
+years <- as.numeric(colnames(all$best_model$data)) # these are the years for the biology model - 1972:2016
+
+# set up climate data
+cli_dat <- data.frame(Year=as.numeric(names(win.sst)),SST=win.sst)
+
+cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))
+
+# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
+
+cli_dat$slope <- 1
+cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
+cli_dat$slope[cli_dat$Year %in% 2014:2019] <- 3
+cli_dat$slope <- as.factor(cli_dat$slope)
+
+# rename rotated
+r <- rotated
+
+Y = rnorm(length(cli_dat$slope))
+cli_dat = as.data.frame(model.matrix(Y ~ Year + SST:slope, data=cli_dat))
+# replace ':' in name for formula
+names(cli_dat) = gsub(":","",names(cli_dat))
+# remove named intercept
+cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
+
+set.seed(99)
+mod = trend_lm(rotated_modelfit = r, all_years = years[years %in% 1981:2019],
+  selected_years = years[years %in% 1981:2019], climate_dat=cli_dat)
+
+names(mod)[2:4] <- c("1981-1988", "1989-2013", "2014-2017")
+summary(mod)
+
+# get 95% CIs
+
+for(i in 2:4){
+x <- sort(mod[,i])
+print(names(mod)[i])
+print(paste("median: ", median(x)))
+print(paste("LCI: ", x[(0.025*25000)]))
+print(paste("UCI: ", x[(0.975*25000)]))
+}
+
+melt.mod <- melt(mod, measure.vars=c("1981-1988", "1989-2013", "2014-2017"))
+names(melt.mod)[3] <- "era"
+melt.mod$covar <- "SST"
+
+# and plot
+# make a SST-only plot
+ggplot(melt.mod, aes(x=value, fill=era)) +
+  theme_bw() + 
+  geom_density(alpha=0.6) +
+  geom_vline(xintercept = 0, lty=2) +
+  scale_fill_manual(values = cb[c(4, 6, 2)]) +
+  theme(legend.title = element_blank(), legend.position = c(0.8, 0.8)) +
+  xlab("Slope") +
+  xlim(-8,7)
+
+ggsave("all plankton era sst slopes.png", width=5, height=4, units="in")
+```

### 2019-11-22 paper draft.Rmd
Export plankton loadings in table

@@
 prop1
+
+# export to share with collaborators
+q.025 <- function(x) {return(quantile(x, probs=0.025))}
+q.975 <- function(x) {return(quantile(x, probs=0.975))}
+export <- data.frame(name=names(loadings.1), mean=colMeans(loadings.1), LCI=apply(loadings.1, 2, q.025), UCI=apply(loadings.1, 2, q.975))
+export[,2:4] <- round(export[,2:4],3)
+export <- export %>%
+  arrange(mean)
+
+write.csv(export, "plankton loadings.csv", row.names = F)
+
 keep <- prop1 > 0.95

### 2019-12-10 paper draft.Rmd
Tweaking figs.

--- 2019-11-19 paper draft.Rmd
@@
 
+# set palette
+cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
 plot.b <- ggplot(plot.trend, aes(year, trend)) +
@@
   geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
-  geom_hline(yintercept = 0) + ylab("Trend value") +
-  theme(axis.title.x = element_blank()) + 
+  geom_hline(yintercept = 0) + ylab("Trend value") + xlab("Year") +
   ggtitle("b) Trend")
@@
   theme_bw() +
-  geom_violin(fill=cb[3]) +
+  geom_violin(fill=cb[3], color=NA) +
   geom_hline(yintercept = 0) +
@@
   coord_flip() +
-  ggtitle("a) Loadings")
+  ggtitle("a) Loadings") +
+  ylim(-2,2)
 

### 2019-12-11 paper draft.Rmd
Changed to 0.9 loadings threshold for climate plots

@@
 
-png("combined GOA climate loading and trend plot.png", 6, 3, units = "in", res=300)
-
-ggpubr::ggarrange(plot.a, plot.b, ncol=2)
-
-dev.off()
-
 # back to the bespoke code!
@@
 prop1
-keep <- prop1 > 0.95
+keep <- prop1 > 0.90 # changing to 0.90 threshold!
 
@@
   ylim(-2,2)
+
+png("combined GOA climate loading and trend plot.png", 6, 3, units = "in", res=300)
+
+ggpubr::ggarrange(plot.a, plot.b, ncol=2)
+
+dev.off()
 
@@
 # trends
-plot.trend <- data.frame(year=1972:2017, trend=as.vector(rotated$trends_mean),
-                         lo=as.vector(rotated$trends_lower),
-                         hi=as.vector(rotated$trends_upper))
-
-plot.b <- ggplot(plot.trend, aes(year, trend)) +
-  theme_bw() +
-  geom_line(color=cb[3]) +
-  geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
-  geom_hline(yintercept = 0) + ylab("Trend value") +
-  theme(axis.title.x = element_blank()) + 
-  ggtitle("b) Trend")
-
-png("combined GOA climate loading and trend plot.png", 8, 4, units = "in", res=300)
-
-ggpubr::ggarrange(plot.a, plot.b, ncol=2)
-
-dev.off()
+# plot.trend <- data.frame(year=1972:2017, trend=as.vector(rotated$trends_mean),
+#                          lo=as.vector(rotated$trends_lower),
+#                          hi=as.vector(rotated$trends_upper))
+# 
+# plot.b <- ggplot(plot.trend, aes(year, trend)) +
+#   theme_bw() +
+#   geom_line(color=cb[3]) +
+#   geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
+#   geom_hline(yintercept = 0) + ylab("Trend value") +
+#   theme(axis.title.x = element_blank()) + 
+#   ggtitle("b) Trend")
+# 
+# png("combined GOA climate loading and trend plot.png", 8, 4, units = "in", res=300)
+# 
+# ggpubr::ggarrange(plot.a, plot.b, ncol=2)
+# 
+# dev.off()
 ```

### 2019-12-11 paper draft.Rmd
Updated plots and loadings threshold

--- 2019-12-10 paper draft.Rmd
@@
 ```{r, echo=F}
-GOA.biol <- readRDS("GOA_biol_9.9.19.rds") 
+GOA.biol <- readRDS("GOA_biol_11.4.19.rds") 
 GOA.biol$summary
@@
 prop1
-keep <- prop1 > 0.95
+keep <- prop1 > 0.9 # changed threshold to 0.9
 
@@
   theme_bw() +
-  geom_violin(fill=cb[3]) +
+  geom_violin(fill=cb[3], color=NA) +
   geom_hline(yintercept = 0) +
   theme(axis.title.y = element_blank()) + ylab("Loading") +
-  coord_flip() +
+  coord_flip() + ylim(-1.5, 1.5) +
   ggtitle("a) Loadings")
@@
 # trends
-plot.trend <- data.frame(year=1972:2017, trend=as.vector(rotated$trends_mean),
+plot.trend <- data.frame(year=1972:2019, trend=as.vector(rotated$trends_mean),
                          lo=as.vector(rotated$trends_lower),
@@
   geom_hline(yintercept = 0) + ylab("Trend value") +
-  theme(axis.title.x = element_blank()) + 
+  xlab("Year") + 
   ggtitle("b) Trend")
 
-png("combined GOA climate loading and trend plot.png", 8, 4, units = "in", res=300)
+png("combined GOA biology loading and trend plot.png", 8, 4, units = "in", res=300)
 
@@
   climate_dat = NULL,
-  trend_number = 1, samples = 50,
+  trend_number = 1, samples = 500,
   ITER = 1000, CHAINS = 1, n_eff = 50, rhat = 1.1, show_plot = TRUE) {
@@
 cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
-cli_dat$slope[cli_dat$Year %in% 2014:2017] <- 3
+cli_dat$slope[cli_dat$Year %in% 2014:2019] <- 3
 cli_dat$slope <- as.factor(cli_dat$slope)
@@
 
-names(mod)[2:4] <- c("1972-1988", "1989-2013", "2014-2017")
+names(mod)[2:4] <- c("1972-1988", "1989-2013", "2014-2019")
 summary(mod)
@@
 
-melt.mod <- melt(mod, measure.vars=c("1972-1988", "1989-2013", "2014-2017"))
+melt.mod <- melt(mod, measure.vars=c("1972-1988", "1989-2013", "2014-2019"))
 names(melt.mod)[3] <- "era"
@@
 ```{r, echo=F}
-
-pdo <- read.csv("pdo.csv")
-
-win.pdo <- pdo$NDJFM
-
-years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2017
-
-# set up climate data
-cli_dat <- data.frame(Year=pdo$YEAR, PDO=win.pdo)
-
-cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(PDO = arm::rescale(PDO))
-
-# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
-
-cli_dat$slope <- 1
-cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
-cli_dat$slope[cli_dat$Year %in% 2014:2017] <- 3
-cli_dat$slope <- as.factor(cli_dat$slope)
-
-# rename rotated
-r <- rotated
-
-Y = rnorm(length(cli_dat$slope))
-cli_dat = as.data.frame(model.matrix(Y ~ Year + PDO:slope, data=cli_dat))
-# replace ':' in name for formula
-names(cli_dat) = gsub(":","",names(cli_dat))
-# remove named intercept
-cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
-
-set.seed(99)
-mod2 = trend_lm(rotated_modelfit = r, all_years = 1972:2017,
-  selected_years = 1972:2016, climate_dat=cli_dat)
-
-names(mod2)[2:4] <- c("1972-1988", "1989-2013", "2014-2017")
-
-summary(mod2)
-
-# get 95% CIs
-
-for(i in 2:4){
-x <- sort(mod2[,i])
-print(names(mod2)[i])
-print(paste("median: ", median(x)))
-print(paste("LCI: ", x[(0.025*25000)]))
-print(paste("UCI: ", x[(0.975*25000)]))
-}
-
-melt.mod2 <- melt(mod2, measure.vars=c("1972-1988", "1989-2013", "2014-2017"))
-names(melt.mod2)[3] <- "era"
-melt.mod2$covar <- "PDO"
+# 
+# pdo <- read.csv("pdo.csv")
+# 
+# win.pdo <- pdo$NDJFM
+# 
+# years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2017
+# 
+# # set up climate data
+# cli_dat <- data.frame(Year=pdo$YEAR, PDO=win.pdo)
+# 
+# cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(PDO = arm::rescale(PDO))
+# 
+# # cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
+# 
+# cli_dat$slope <- 1
+# cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
+# cli_dat$slope[cli_dat$Year %in% 2014:2017] <- 3
+# cli_dat$slope <- as.factor(cli_dat$slope)
+# 
+# # rename rotated
+# r <- rotated
+# 
+# Y = rnorm(length(cli_dat$slope))
+# cli_dat = as.data.frame(model.matrix(Y ~ Year + PDO:slope, data=cli_dat))
+# # replace ':' in name for formula
+# names(cli_dat) = gsub(":","",names(cli_dat))
+# # remove named intercept
+# cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
+# 
+# set.seed(99)
+# mod2 = trend_lm(rotated_modelfit = r, all_years = 1972:2017,
+#   selected_years = 1972:2016, climate_dat=cli_dat)
+# 
+# names(mod2)[2:4] <- c("1972-1988", "1989-2013", "2014-2017")
+# 
+# summary(mod2)
+# 
+# # get 95% CIs
+# 
+# for(i in 2:4){
+# x <- sort(mod2[,i])
+# print(names(mod2)[i])
+# print(paste("median: ", median(x)))
+# print(paste("LCI: ", x[(0.025*25000)]))
+# print(paste("UCI: ", x[(0.975*25000)]))
+# }
+# 
+# 
+# melt.mod2 <- melt(mod2, measure.vars=c("1972-1988", "1989-2013", "2014-2017"))
+# names(melt.mod2)[3] <- "era"
+# melt.mod2$covar <- "PDO"
 
@@
   theme(legend.title = element_blank(), legend.position = c(0.8, 0.8)) +
-  xlab("Slope") +
+  xlab("Slope") + ylab("Density") +
   xlim(-5,12)

### 2019-12-11 paper draft.Rmd
Typo
Updated plots for ms. revisions

@@
 
-clim.regimes <- data.frame(year=1950:2019, probability=kk$data$median)
+clim.regimes <- data.frame(year=1950:2019, probability=kk$data$median, lwr=kk$data$lwr, upr=kk$data$upr)
 
@@
 
-biol.regimes <- data.frame(year=1972:2017, probability=kk$data$median)
-
-biol.regimes$data.set <- "Biology 1972-2017"
+biol.regimes <- data.frame(year=1972:2019, probability=kk$data$median, lwr=kk$data$lwr, upr=kk$data$upr)
+
+biol.regimes$data.set <- "Biology 1972-2019"
 
@@
 We're interested in testing the hypothesis that the slope on sst was greater during 1972:1988 (high Aleutian Low variance) than during 1989:2013 (low Aleutian Low variance).
-Also interested in asking which of these periods the 2014-106 heat wave years are most similar to.
+Also interested in asking which of these periods the 2014-2016 heat wave years are most similar to.
 
@@
 ```{r, echo=F}
-GOA.ich <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_ich_7.19.18.one.trend.rds") 
+# GOA.ich <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_ich_7.19.18.one.trend.rds") 
+GOA.ich <- readRDS("/Users/MikeLitzow/Documents/github/ewi-in-recovery/ewi-in-recovery/all_plankton_11.20.19.one.trend.rds") 
 GOA.ich$summary
@@
 # load updated ewidata!
-ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
-
-# limit to GOA
-GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
-unique(GOA$code)
-
-# limit to ichthy
-
-keep <- grep("ICH", GOA$code)
-
-GOA <- GOA[keep,]
+# ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
+# 
+# # limit to GOA
+# GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
+# unique(GOA$code)
+# 
+# # limit to ichthy
+# 
+# keep <- grep("ICH", GOA$code)
+# 
+# GOA <- GOA[keep,]
 
 # recover the years and names for plots!
+# using GOA from 'updated plankton model.R'
 sub_data = GOA
@@
 rotated = rotate_trends(GOA.ich$best_model) 
-new.names <- c("sandlance", "ronquils", "walleye.pollock", "Pacific.cod", "flathead.sole", "n.rock.sole", "rockfish")
-
- check <- data.frame(names=names, newnames=new.names)
- check
+# new.names <- c("sandlance", "ronquils", "walleye.pollock", "Pacific.cod", "flathead.sole", "n.rock.sole", "rockfish")
+# 
+#  check <- data.frame(names=names, newnames=new.names)
+#  check
 
@@
 set.seed(11)
-f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)
+f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 5)
 
@@
 
-ichthyo.regimes <- data.frame(year=plot.trend$year, probability=kk$data$median)
-
-ichthyo.regimes$data.set <- "Ichthyoplankton 1980-2015"
+ichthyo.regimes <- data.frame(year=plot.trend$year, probability=kk$data$median, lwr=kk$data$lwr, upr=kk$data$upr)
+
+ichthyo.regimes$data.set <- "Plankton 1981-2019"
 
 plot.regimes <- rbind(clim.regimes, biol.regimes, ichthyo.regimes)
-
-ggplot(plot.regimes, aes(year, probability, color=data.set)) +
-  theme_bw() +
-  geom_line()
-
+plot.regimes$jitter <- ifelse(plot.regimes$data.set=="Biology 1972-2019" & plot.regimes$probability>0.99, 1.002, plot.regimes$probability)
+
+plot.regimes$order <- ifelse(plot.regimes$data.set=="Plankton 1981-2019", 1,
+                             ifelse(plot.regimes$data.set=="Biology 1972-2019", 2, 3))
+
+plot.regimes$data.set <- reorder(plot.regimes$data.set, plot.regimes$order)
+
+ggplot(plot.regimes, aes(year, jitter, color=data.set, fill=data.set)) +
+  theme_bw() +
+  geom_line() +
+  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.2, color=NA) +
+  scale_color_manual(values=cb[c(2,7,6)], guide=guide_legend(reverse=T)) + scale_fill_manual(values=cb[c(2,7,6)], guide=guide_legend(reverse=T)) +
+  xlab("Year") + ylab("Probability") +
+  theme(legend.title = element_blank(), legend.position = 'top')
+
+ggsave("updated HMM regime plot.png", width = 6, height = 4, units='in)
 ```

### 2019-12-11 updated plankton model.R
Updated for single plankton model

@@ -140,10 +140,11 @@ GOA <- ewidata
 drop <- grep("ICY.ZOOP.TOTDEN", GOA$code)
 GOA <- GOA[-drop,]
 
-# select ICY
-keep <- grep("ICY", GOA$code)
+# select ICY, PP, Seward, CPR, and ichthyo!
+keep <- c(grep("ICY", GOA$code), grep("PP_", GOA$code), grep("CPR", GOA$code), grep("SEWAR", GOA$code), grep("ICH", GOA$code))
 GOA <- GOA[keep,]
-GOA$value <- log(GOA$value)
+
+# GOA$value <- log(GOA$value)
 # # and drop ichthyoplankton
 # drop <- grep("ICH", GOA$code)
 # GOA <- GOA[-drop,]
@@ -181,13 +182,14 @@ ggplot(GOA, aes(value)) +
   geom_histogram(fill="grey", color="black", bins=10) + 
   facet_wrap(~code, scales="free")
 
-trns <- grep("amp", GOA$code)
+# log transform icy, seward, and bloom amplitude
+trns <- c(grep("amp", GOA$code), grep("ICY", GOA$code), grep("SEWA", GOA$code))
 GOA$value[trns] <- log(GOA$value[trns])
 
 sub_data = GOA
 
 # set the name for the model output!
-name <- "ICY_11.20.19.one.trend"
+name <- "all_plankton_11.20.19.one.trend"
 
 
 # and now run DFA!

### 2019-12-11 ewidata update - GOA.R
Meaningless

@@ -611,6 +611,7 @@ sub_data <- goa.clim %>%
   filter(year >= 2005) %>%
   gather(key="code", value="value", -year)
 
+
 # ONE! trend model
 # with original data
 max_trends = 1

### 2019-12-11 paper draft.Rmd
Little tweak
Changed jitter

@@
 plot.regimes <- rbind(clim.regimes, biol.regimes, ichthyo.regimes)
-plot.regimes$jitter <- ifelse(plot.regimes$data.set=="Biology 1972-2019" & plot.regimes$probability>0.99, 1.002, plot.regimes$probability)
+plot.regimes$jitter <- ifelse(plot.regimes$data.set=="Biology 1972-2019" & plot.regimes$probability>0.99, 1.004, plot.regimes$probability)
 
@@
 
-ggsave("updated HMM regime plot.png", width = 6, height = 4, units='in)
+ggsave("updated HMM regime plot.png", width = 6, height = 4, units='in')
 ```

### 2019-12-13 paper draft.Rmd
Added unified plankton model plots

@@
 # GOA.ich <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_ich_7.19.18.one.trend.rds") 
-GOA.ich <- readRDS("/Users/MikeLitzow/Documents/github/ewi-in-recovery/ewi-in-recovery/all_plankton_11.20.19.one.trend.rds") 
+GOA.ich <- readRDS("all_plankton_12.11.19.one.trend.rds") 
 GOA.ich$summary
@@
 ```{r, echo=F}
-all <- readRDS("all_plankton_11.20.19.one.trend.rds") 
+all <- readRDS("all_plankton_12.11.19.one.trend.rds") 
 all$summary
@@
 
-keep <- prop1 > 0.95
+keep <- prop1 > 0.90 # changing to 0.9!
 
@@
   theme(title = element_text(size=8)) +
-  ylim(-2.5,2.5)
+  ylim(-2,2)
 
@@
   ggtitle(("c) Plankton slope on SST")) +
-  theme(title = element_text(size=8))
+  theme(title = element_text(size=8)) +
+  xlim(-3.2,0.5)
 
@@
 ```{r}
-png("combined all plankton time series plots.png", 8, 2.5, units="in", res=300)
+png("combined all plankton time series plots.png", 10, 4, units="in", res=300)
 
 ggpubr::ggarrange(plot.a, plot.b, plot.c,
-          ncol=3, widths=c(1, 1, 0.8), nrow=1)
+          ncol=3, widths=c(1, 1.3, 0.8), nrow=1)
 

### 2019-12-13 correlation.R
Plot tweaks

@@ -34,8 +34,8 @@ Y <- as.matrix(Y[, -which(names(Y) == "code")])
 
 # set new names
 new.names <- c("East.spring.SST", "East.winter.SST", "SLP.gradient",
-  "Papa.advection", "GAK1.salinity", "West.spring.SST", "West.winter.SST",
-  "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149")
+               "Papa.advection", "GAK1.salinity", "West.spring.SST", "West.winter.SST",
+               "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149")
 
 # look at rolling correlations between fitted and observed values, as in Ecology paper
 # using 25-yr rolling windows!
@@ -53,9 +53,9 @@ fit_cor_stan <- function(.x) {
   ))
   stan_dat <- list(x = cbind(.x$trend_value, .x$data_value), N = nrow(.x))
   m <- sampling(stan_correlation_m,
-    data = stan_dat,
-    iter = 500, chains = 1, verbose = FALSE, refresh = 0,
-    seed = 1947823,
+                data = stan_dat,
+                iter = 500, chains = 1, verbose = FALSE, refresh = 0,
+                seed = 1947823,
   )
   rho_post <- rstan::extract(m)$rho
   tibble(
@@ -78,7 +78,7 @@ orig_data <- reshape2::melt(modelfit$data) %>%
   as_tibble() %>%
   rename(variable = Var1, year = Var2, data_value = value)
 orig_and_pred <- inner_join(pred_long, orig_data,
-  by = c("year", "variable")
+                            by = c("year", "variable")
 ) %>%
   filter(!is.na(data_value))
 
@@ -120,23 +120,23 @@ out %>%
     upr2 = quantile(rho_post, probs = 0.75),
     med = quantile(rho_post, probs = 0.5)
   ) %>%
-  # filter(!(middle < 1970 & var_name == "GAK1 salinity")) %>%
+  filter(!(middle < 1970 & var_name == "GAK1 salinity")) %>% # renoving data-poor GAK1 windows
   ggplot(aes(x = middle, y = med, ymin = lwr, ymax = upr)) +
-  geom_hline(yintercept = 0, lty = 2, alpha = 0.2) +
+  # geom_hline(yintercept = 0, lty = 2, alpha = 0.2) +
[…]

### 2019-12-13 updated plankton model.R
Updated unified plankton model

@@ -11,8 +11,8 @@ library(gtools)
 library(bayesdfa)
 
 # load updated ewidata!
-ewidata <- read.csv("ewidata.csv", row.names=1)
-
+# ewidata <- read.csv("ewidata.csv", row.names=1)
+#
 # now load sewardline updates
 
 sew <- read.csv("sewardline.updates.csv")
@@ -63,7 +63,7 @@ ggplot(pp, aes(value)) +
   geom_histogram(bins=8) +
   facet_wrap(~code, scales="free")
 
-# I think we can live with that distribution 
+# I think we can live with that distribution
 # - although I should ask Eric/Sean if the bimodal distribution in duration is a problem for the DFA models
 
 pp$code <- paste("PP_", pp$code, sep="")
@@ -72,7 +72,7 @@ pp$system <- "EGOA"
 pp$subtype <- NA
 
 ewidata <- rbind(ewidata, pp)
-# 
+#
 # # now updated CPR data
 
 cpr <- read.csv("CPR.2019.csv")
@@ -125,12 +125,12 @@ cor(check[,4:9])
 
 # combine with ewidata
 ewidata <- rbind(ewidata, icy)
-
-# and save
-write.csv(ewidata, "/Users/MikeLitzow/Documents/R/FATE/ewidata/inst/extdata/ewidata_11.20.19", row.names = F)
+#
+# # and save
+# write.csv(ewidata, "/Users/MikeLitzow/Documents/R/FATE/ewidata/inst/extdata/ewidata_11.20.19", row.names = F)
 write.csv(ewidata, "ewidata.plankton.updated.csv", row.names=F)
 
-
+ewidata <- read.csv("ewidata_11.20.19.csv")
 # GoA biology
 
 codes <- unique(ewidata$code)
@@ -148,38 +148,38 @@ GOA <- GOA[keep,]
 # # and drop ichthyoplankton
 # drop <- grep("ICH", GOA$code)
 # GOA <- GOA[-drop,]
-# 
-# # # and try dropping PP duration, as this is bimodal and might give trouble with the model 
+#
+# # # and try dropping PP duration, as this is bimodal and might give trouble with the model
 # # # returning bimodal loadings
 # # drop <- grep("_dur", GOA$code)
 # # GOA <- GOA[-drop,]
-# 
+#
 # unique(GOA$code)
-# 
+#
 # # now check for outliers!
-# ggplot(GOA, aes(x=year, y = value))  + 
[…]

### 2019-12-15 paper draft.Rmd
Calculated proportion of loadings above and below 0

@@
 
+# and look at the proportion of posteriors > or < 0 for each era
+for(i in 2:4){
+x <- mod[,i]
+print(names(mod)[i])
+print(paste("Proportion < 0: ", sum(x < 0)/length(x)))
+print(paste("Proportion > 0: ", sum(x > 0)/length(x)))
+}
+
 melt.mod <- melt(mod, measure.vars=c("1972-1988", "1989-2013", "2014-2019"))
@@
 ```{r, echo=F}
-all <- readRDS("all_plankton_12.11.19.one.trend.rds") 
+# all <- readRDS("all_plankton_12.11.19.one.trend.rds") 
+all <- readRDS("GOA_plankton_11.18.19.one.trends.rds") 
 all$summary
@@
 print(paste("UCI: ", x[(0.975*25000)]))
+
+# and look at the proportion of posteriors > or < 0
+
+x <- mod[,2]
+print(names(mod)[i])
+print(paste("Proportion < 0: ", sum(x < 0)/length(x)))
+print(paste("Proportion > 0: ", sum(x > 0)/length(x)))
 

### 2019-12-16 ewidata update - GOA.R
Changed 1st 15-yr period for DFA to 1972-1986, added violoiun annd trend plots of same

@@ -10,7 +10,7 @@ dat <- read.csv("ewidata.csv", row.names = 1)
 
 levels(dat$system)
 
-# restrict to GOA 
+# restrict to GOA
 
 goa.dat <- dat %>%
   filter(system %in% c("EGOA", "WGOA"))
@@ -19,7 +19,7 @@ ggplot(goa.dat, aes(year, value)) +
   theme_bw() +
   geom_line() +
   geom_point() +
-  facet_wrap(~code, scales="free_y") 
+  facet_wrap(~code, scales="free_y")
 
 # extend SST data
 # load ERSSTv5 data for the N. Pacific
@@ -54,14 +54,14 @@ f2 <- function(x) tapply(x, m, sd)
 SST <- ncvar_get(nc,  "sst")
 # Change data from a 3-D array to a matrix of monthly data by grid point:
 # First, reverse order of dimensions ("transpose" array)
-SST <- aperm(SST, 3:1)  
+SST <- aperm(SST, 3:1)
 
 # Change to matrix with column for each grid point, rows for monthly means
-SST <- matrix(SST, nrow=dim(SST)[1], ncol=prod(dim(SST)[2:3]))  
+SST <- matrix(SST, nrow=dim(SST)[1], ncol=prod(dim(SST)[2:3]))
 
 # Keep track of corresponding latitudes and longitudes of each column:
-sst.lat <- rep(sst.y, length(sst.x))   
-sst.lon <- rep(sst.x, each = length(sst.y))   
+sst.lat <- rep(sst.y, length(sst.x))
+sst.lon <- rep(sst.x, each = length(sst.y))
 dimnames(SST) <- list(as.character(sst.d), paste("N", sst.lat, "E", sst.lon, sep=""))
 
 ##############################################################################
@@ -83,14 +83,14 @@ SST[,blank] <- NA
 SST.mean <- colMeans(SST)
 z <- t(matrix(SST.mean,length(sst.y)))  # Re-shape to a matrix with latitudes in columns, longitudes in rows
 image(sst.x,sst.y,z, col=tim.colors(64), xlim=c(160,240), ylim=c(40,66))
-contour(sst.x, sst.y, z, add=T)  
+contour(sst.x, sst.y, z, add=T)
 map('world2Hires',fill=F,xlim=c(130,250), ylim=c(20,66),add=T, lwd=2)
 
 # looks good!
 
-# now remove seasonal signal and scale 
+# now remove seasonal signal and scale
[…]

### 2019-12-30 ewidata update - GOA.R
Changed early era

@@ -699,7 +699,7 @@ rotated.2 = rotate_trends(GOA.clim.era2$best_model)
 loadings.2 <- as.data.frame(rotated.2$Z_rot[,,1])
 names(loadings.2)  <- new.names
 
-loadings.1$era <- "1950-1964"
+loadings.1$era <- "1972-1986"
 loadings.2$era <- "2005-2019"
 
 

### 2019-12-30 ewidata update - GOA.R
Improved era plot

@@ -708,21 +708,25 @@ plot.load <- rbind(loadings.1, loadings.2) %>%
   gather(key, value, -era)
 
 
-rank <- tapply(plot.load$value, plot.load$key, mean, na.rm=T)
+rank <- tapply(plot.load$value[plot.load$era=="1972-1986"], plot.load$key[plot.load$era=="1972-1986"], median, na.rm=T)
 plot.load$rank <- rank[match(plot.load$key, names(rank))]
 plot.load$key <- reorder(plot.load$key, -plot.load$rank)
+
+# and order eras!
+plot.load$era.order <- as.factor(ifelse(plot.load$era=="1972-1986", 2, 1))
 
 # set pallette
 cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 
-plot.a <- ggplot(plot.load, aes(key, value, fill=era)) +
+plot.a <- ggplot(plot.load, aes(key, value, fill=era.order)) +
   geom_violin(color=NA) +
-  scale_fill_manual(values=cb[c(2,4)]) +
+  scale_fill_manual(values=cb[c(4,2)], labels=c("2005-2019", "1972-1986")) +
   theme_bw() +
   ylim(-2, 2) + geom_hline(yintercept = 0, lty=2) +
   theme(legend.title = element_blank(), axis.title.y = element_blank(), legend.position = 'top',
         title = element_text(size=9)) +
-  coord_flip() + ylab("Loading") + ggtitle("a) Era-specific loadings")
+  coord_flip() + ylab("Loading") + ggtitle("a) Era-specific loadings") +
+  guides(fill=guide_legend(reverse=T))
 
 plot.trend <- data.frame(year=1972:1986, trend=as.vector(rotated.1$trends_mean),
                          lo=as.vector(rotated.1$trends_lower),

### 2019-12-31 ewidata update - GOA biology.R
Fixed plot

@@ -283,12 +283,12 @@ dev.off()
 # last 15
 sub_data <- GOA %>%
   select(year, code, value) %>%
-  filter(year >= 2003)
+  filter(year >= 2005)
 
 # ONE! trend model
 # with original data
 max_trends = 1
-name <- "GOA_biol_1_trend_original_data_2003_2017"
+name <- "GOA_biol_1_trend_original_data_2005_2019"
 
 # reshape data
 melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
@@ -384,21 +384,29 @@ loadings.2$era <- "2005-2019"
 plot.load <- rbind(loadings.1, loadings.2) %>%
   gather(key, value, -era)
 
-rank <- tapply(plot.load$value, plot.load$key, mean)
+# check!
+plot(plot.load$value[plot.load$era=="1972-1986"], plot.load$value[plot.load$era=="2005-2019"])
+
+
+rank <- tapply(plot.load$value[plot.load$era=="1972-1986"], plot.load$key[plot.load$era=="1972-1986"], median, na.rm=T)
 plot.load$rank <- rank[match(plot.load$key, names(rank))]
 plot.load$key <- reorder(plot.load$key, -plot.load$rank)
 
-# set pallette
+# and order eras!
+plot.load$era.order <- as.factor(ifelse(plot.load$era=="1972-1986", 2, 1))
+
+# set palette
 cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 
-plot.a <- ggplot(plot.load, aes(key, value, fill=era)) +
+plot.a <- ggplot(plot.load, aes(key, value, fill=era.order)) +
   geom_violin(color=NA) +
-  scale_fill_manual(values=cb[c(2,4)]) +
+  scale_fill_manual(values=cb[c(4,2)], labels=c("2005-2019", "1972-1986")) +
   theme_bw() +
-  ylim(-1.8, 2) + geom_hline(yintercept = 0, lty=2) +
+  ylim(-2, 2) + geom_hline(yintercept = 0, lty=2) +
   theme(legend.title = element_blank(), axis.title.y = element_blank(), legend.position = 'top',
         title = element_text(size=9)) +
-  coord_flip() + ylab("Loading") + ggtitle("a) Era-specific loadings")
+  coord_flip() + ylab("Loading") + ggtitle("a) Era-specific loadings") +
[…]

### 2019-12-31 paper draft.Rmd
New plankton DFA plot
Tweaks to plots

@@
 ```{r, echo=F}
-# all <- readRDS("all_plankton_12.11.19.one.trend.rds") 
-all <- readRDS("GOA_plankton_11.18.19.one.trends.rds") 
+all <- readRDS("all_plankton_12.11.19.one.trend.rds")
+# all <- readRDS("GOA_plankton_11.18.19.one.trends.rds") 
 all$summary
@@
   theme_bw() +
-  geom_density(fill=cb[3]) + 
+  geom_density(fill=cb[3], color=NA) + 
   geom_hline(yintercept=0) + 
@@
 dev.off()
+#######
+
+plot.null <- ggplot() + theme_void()
+
+png("combined all plankton time series plots.png", 6, 6, units="in", res=300)
+
+ggpubr::ggarrange(ggpubr::ggarrange(plot.null,plot.a, plot.null, ncol=3, nrow=1, widths=c(0.1, 1, 0.2)),
+                  ggpubr::ggarrange(plot.b, plot.c, ncol=2, widths=c(1.5, 0.8), nrow=1),
+                  nrow=2, heights=c(1, 0.8))
+
+dev.off()
+
 ```

### 2019-12-31 ewidata update - GOA biology.R
Plot bug fix!

@@ -367,13 +367,13 @@ check <- cbind(as.character(names), new.names)
 check
 
 # back to the bespoke code!
-loadings.1 <- as.data.frame(rotated$Z_rot[,,1])
+loadings.1 <- as.data.frame(rotated.1$Z_rot[,,1])
 names(loadings.1)  <- new.names
 
 # now....era2
 GOA.biol.era2 <- readRDS("GOA_biol_1_trend_original_data_2005_2019.rds") # read in GOA climate model
 rotated.2 = rotate_trends(GOA.biol.era2$best_model)
-loadings.2 <- as.data.frame(rotated$Z_rot[,,1])
+loadings.2 <- as.data.frame(rotated.2$Z_rot[,,1])
 names(loadings.2)  <- new.names
 
 loadings.1$era <- "1972-1986"

### 2020-01-03 paper draft.Rmd
Tweaked climate plot labels

@@
 # set new names
-new.names <- c("East.spring.SST", "East.winter.SST", "SLP.gradient", "Papa.advection", "GAK1.salinity", "West.spring.SST", "West.winter.SST", "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149")
+new.names <- c("East spring SST", "East winter SST", "SLP gradient", "Papa advection", "GAK1 salinity", "West spring SST", "West winter SST", " Upwell 54 134", " Upwell 57 137", " Upwell 60 146", " Upwell 60 149")
 
@@
 
-# set new names
-new.names <- c("East.spring.SST", "East.winter.SST", "SLP.gradient", "Papa.advection", "GAK1.salinity", "West.spring.SST", "West.winter.SST", "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149")
-
 plot.trend <- data.frame(year=1950:2019, trend=as.vector(rotated$trends_mean),
@@
 prop1
-keep <- prop1 > 0.95
+keep <- prop1 > 0.90
 

### 2020-01-03 ewidata update - GOA.R
Tweaked climate labels for plot

@@ -687,7 +687,7 @@ rotated.1 = rotate_trends(GOA.clim.era1$best_model)
 
 # set new names
 new.names <- c("East spring SST", "East winter SST", "SLP gradient", "Papa advection",
-               "GAK1 salinity", "West spring SST", "West winter SST", "Downwell 54 134", "Downwell 57 137", "Downwell 60 146", "Downwell 60 149")
+               "GAK1 salinity", "West spring SST", "West winter SST", "Upwell 54 134", "Upwell 57 137", "Upwell 60 146", "Upwell 60 149")
 
 # back to the bespoke code!
 loadings.1 <- as.data.frame(rotated.1$Z_rot[,,1])

### 2020-01-03 correlation.R
Fixed upwelling labels

@@ -35,7 +35,7 @@ Y <- as.matrix(Y[, -which(names(Y) == "code")])
 # set new names
 new.names <- c("East.spring.SST", "East.winter.SST", "SLP.gradient",
                "Papa.advection", "GAK1.salinity", "West.spring.SST", "West.winter.SST",
-               "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149")
+               "Upwelling.54.134", "Upwelling.57.137", "Upwelling.60.146", "Upwelling.60.149")
 
 # look at rolling correlations between fitted and observed values, as in Ecology paper
 # using 25-yr rolling windows!

### 2020-01-03 paper draft.Rmd
Fixed biology plot labels

@@
 
-new.names <- c("Chignik.coho", "Chignik.pink", "Chiniak.eulachon", "Chiniak.jellyfish", "Chiniak.P.cod", "Chiniak.shrimp", "Cook.Inlet.coho", "Kodiak.coho", "Kodiak.pink", "Pavlof.capelin", "Pavlof.eulachon", "Pavlof.jellyfish", "Pavlof.P.cod", "Pavlof.shrimp", "Prince.William.coho", "Southeast.coho", "Southeast.herring", "Southeast.pink", "South.Peninsula.coho", "South.Peninsula.pink")
+new.names <- c("Chignik coho", "Chignik pink", "Chiniak eulachon", "Chiniak jellyfish", "Chiniak P.
+cod", "Chiniak shrimp", "Cook Inlet coho", "Kodiak coho", "Kodiak pink", "Pavlof capelin", "Pavlof eulachon", "Pavlof jellyfish", "Pavlof P.
+cod", "Pavlof shrimp", "Prince William coho", "Southeast coho", "Southeast herring", "Southeast pink", "South Peninsula coho", "South Peninsula pink")
 
